title: "Gramminoid Composition Analysis"
author: "Courtney Buchanan"
date: "2024-09-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#library packages
```{r}

library(ggplot2)
library(tidyverse)
library(dplyr)
#install.packages("FSA")
library(FSA)
library(tidyr)
#install.packages("gggpubr")
library(ggpubr)
library(MKinfer)
library(car)

```
#read in files
```{r}

plant_family<-read.csv("./9_10_24plant_family_percents_outputoftrnlcleaned.csv") #diet composition from TRNL file output
metadata<-read.csv("./11_9_23_metadata_no_copy.csv") #sample metadata
individual_BCS<-read.csv("./bcs_by_observation_5_24_23.csv") #body condition observations
HMA_RAP<-read.csv("./allHMAs_withRAP_4_18_24.csv") # RAP biomass and cover data for all HMAs

metadata_with_RAP<-read.csv("./5_9_23_metadata_no_copy_withRAP.csv")#read in buffered RAP data with biomass and cover data buffered for each sample


```

#prep dataframe for analysis
```{r}
plant_family1<-as.data.frame(plant_family) #make a dataframe to edit from original 
plant_family1<-tibble::column_to_rownames((as.data.frame(plant_family1)), "X") #puts sample name as the row name

colnames(plant_family1)<-as.character(plant_family1[498,]) #takes family name and makes it the column names 

plant_family1<-plant_family1[2:493,]
```

##join metadata
```{r}
#make a sample column to join by 

plant_family1<- tibble::rownames_to_column((as.data.frame(plant_family1)), "sample") #makes "sample" a column to join with metadata

plant_family_meta<- left_join(plant_family1, metadata, by="sample") # join diet data with horse sample metadata

plant_family_meta_wild<- subset(plant_family_meta, domestication == "wild")#get rid of domestic for now

```

#group for where I want to use HMA as unit and for graphs
```{r}
#make the columns I want to use numeric
plant_family_meta_wild$Poaceae<-as.numeric(plant_family_meta_wild$Poaceae)
plant_family_meta_wild$Cyperaceae<-as.numeric(plant_family_meta_wild$Cyperaceae)
plant_family_meta_wild$Juncaceae<-as.numeric(plant_family_meta_wild$Juncaceae)

plant_family_meta_wild$Poales<-plant_family_meta_wild$Poaceae + plant_family_meta_wild$Cyperaceae + plant_family_meta_wild$Juncaceae #this combines poaceae, cyperaceae, and Juncaceae into one grasslike column called "Poales"

# Use the amount Poales to summarize 

#Somehow we got 16 samples for WMP so exclude WMP11 which has least fresh fecal score
plant_family_meta_wild<-subset(plant_family_meta_wild, sample != "WMP11")

# this leaves each hma season combo as its own
plant_order_by_HMA_season<-plant_family_meta_wild%>% #Looking at average for Poales only    
  group_by( study.area, area_season, season )%>%
  summarise(mean_Poales=mean(Poales), sd=sd(Poales), number= n())

View(plant_order_by_HMA_season)

```
#find average graminoids per season
```{r}
graminoids_by_HMA_summer<- subset(plant_order_by_HMA_season, season == "summer") #subset summer
mean(graminoids_by_HMA_summer$mean_Poales)# 58.102 summer average including SB
mean(subset(graminoids_by_HMA_summer, study.area != "SB")$mean_Poales) # 59.43886 summer average no SB 

sd(graminoids_by_HMA_summer$mean_Poales)# 18.16 including SB
sd(subset(graminoids_by_HMA_summer, study.area != "SB")$mean_Poales) # 17.9647 no SB 



graminoids_by_HMA_winter<- subset(plant_order_by_HMA_season, season == "winter")#subset winter
mean(graminoids_by_HMA_winter$mean_Poales)# 49.874 winter average
sd(graminoids_by_HMA_winter$mean_Poales)# 17.74981 winter sd

mean(plant_order_by_HMA_season$mean_Poales) #overall average = 54.12116 with SB
mean(subset(plant_order_by_HMA_season, study.area != "SB")$mean_Poales) # 54.6566 average no SB 
sd(subset(plant_order_by_HMA_season, study.area != "SB")$mean_Poales) # 18.20867 no SB
sd(plant_order_by_HMA_season$mean_Poales) #18.14916 with SB
```
#subset to each study area for analyses

```{r}

#subset to HMA so I can do one study area at a time in the analysis
plant_order_AT<-subset(plant_family_meta_wild, study.area == "AT")
plant_order_CN<-subset(plant_family_meta_wild, study.area == "CN")
plant_order_FR<-subset(plant_family_meta_wild, study.area == "FR")
plant_order_GM<-subset(plant_family_meta_wild, study.area == "GM")
plant_order_LBC<-subset(plant_family_meta_wild, study.area == "LBC")
plant_order_MP<-subset(plant_family_meta_wild, study.area == "MP")
plant_order_NH<-subset(plant_family_meta_wild, study.area == "NH")
plant_order_ON<-subset(plant_family_meta_wild, study.area == "ON")
plant_order_PB<-subset(plant_family_meta_wild, study.area == "PB")
plant_order_PN<-subset(plant_family_meta_wild, study.area == "PN")
plant_order_RR<-subset(plant_family_meta_wild, study.area == "RR")
plant_order_SC<-subset(plant_family_meta_wild, study.area == "SC")
plant_order_SS<-subset(plant_family_meta_wild, study.area == "SS")
plant_order_SW<-subset(plant_family_meta_wild, study.area == "SW")
plant_order_TP<-subset(plant_family_meta_wild, study.area == "TP")

```

#Seasonal comparison
## T-tests
###range
```{r}
## t test- not paired 
t.test(mean_Poales ~ season, data= plant_order_by_HMA_season) #not significant p=0.2101


#t-test paired...
View(plant_order_by_HMA_season)
range_sum<- plant_order_by_HMA_season[plant_order_by_HMA_season$season == "summer",] #pull summer HMA averages
range_sum<- range_sum[range_sum$study.area != "SB",]#exclude SB because no winter sample
range_sum<- range_sum$mean_Poales #take only the graminoid column
range_sum # View values
range_win<- plant_order_by_HMA_season[plant_order_by_HMA_season$season == "winter",]#pull winter HMA averages 
range_win<-range_win$mean_Poales #take only the graminoid column
range_win #view values


t.test(x=range_sum, y=range_win, paired=TRUE, var.equal = T) #when they are paired we have a significant p-value 0.03617


#look at histogram and differences 
hist(range_sum, breaks=10) #summer- not normal
hist(range_win, breaks= 10) #winter- not normal, but not terrible
hist(plant_order_by_HMA_season$mean_Poales)#all seasons- does not look super normal...
#for a paired t-test it is the difference that needs to be normal?
hist(range_sum - range_win, breaks = 10) #still not normal... 


boxplot(plant_order_by_HMA_season$mean_Poales ~ plant_order_by_HMA_season$season) #likely not equal, summer maybe slight more variable- actual numbers close
var(range_win) # 0.0315
var(range_sum) # 0.0323


```


###individual HMAs
```{r}
#for each study area do a 2 sample t test independent, then pair for the max and minimum variability to compare the p-values. 

#AT
t.test(Poales ~ season, data= plant_order_AT) #not significant p=0.8571

AT_sum<- plant_order_AT[plant_order_AT$season == "summer",]
AT_sum<-AT_sum$Poales
AT_win<- plant_order_AT[plant_order_AT$season == "winter",]
AT_win<-AT_win$Poales

#min variation
t.test(x=sort(AT_sum), y=sort(AT_win,decreasing=FALSE), paired=TRUE) # not significant p = 0.7315
#max variation
t.test(x=sort(AT_sum), y=sort(AT_win,decreasing=TRUE), paired=TRUE)# not significant p = 0.8848
#not paired
t.test(x=AT_sum, y = AT_win, paired= FALSE) #not significant p=0.8571
t.test(x=AT_sum, y = AT_win, paired= FALSE, var.equal = TRUE) #not significant p=0.8566

#histograms and assumptions
hist(plant_order_AT$Poales, breaks = 10) #all observations, not normal
hist(AT_sum, breaks = 10)#summer, not normal
hist(AT_win, breaks = 10) #winter, not normal
hist(AT_sum - AT_win, breaks = 10) #difference, kinda normal 

boxplot(plant_order_AT$Poales ~ plant_order_AT$season) #likely not equal, winter more variable, but summer has outlier 
var(AT_win) #0.12
var(AT_sum) #0.03


#CN
t.test(Poales ~ season, data= plant_order_CN) #not significant p=0.1975

CN_sum<- plant_order_CN[plant_order_CN$season == "summer",]
CN_sum<-CN_sum$Poales
CN_win<- plant_order_CN[plant_order_CN$season == "winter",]
CN_win<-CN_win$Poales

#min variation
t.test(x=sort(CN_sum), y=sort(CN_win,decreasing=FALSE), paired=TRUE) # significant p = 0.0305
#max variation
t.test(x=sort(CN_sum), y=sort(CN_win,decreasing=TRUE), paired=TRUE)# not significant p = 0.3133
#not paired
t.test(x=CN_sum, y = CN_win, paired= FALSE) #not significant p=0.1975
t.test(x=CN_sum, y = CN_win, paired= FALSE, var.equal = TRUE) #not significant p=0.1931

#histograms
hist(plant_order_CN$Poales, breaks = 10) #all observations, not normal
hist(CN_sum, breaks = 10)#summer, not normal
hist(CN_win, breaks = 10) #winter, not normal
hist(CN_sum - CN_win, breaks = 10) #difference, not normal  


boxplot(plant_order_CN$Poales ~ plant_order_CN$season) #likely not equal, winter more variable, summer outlier
var(CN_win) # 0.15
var(CN_sum) # 0.03


#FR
t.test(Poales ~ season, data= plant_order_FR)#significant p= 0.0002

FR_sum<- plant_order_FR[plant_order_FR$season == "summer",]
FR_sum<-FR_sum$Poales
FR_win<- plant_order_FR[plant_order_FR$season == "winter",]
FR_win<-FR_win$Poales

#min variation
t.test(x=sort(FR_sum), y=sort(FR_win,decreasing=FALSE), paired=TRUE) #significant p= 4.8 e-10
#max variation
t.test(x=sort(FR_sum), y=sort(FR_win,decreasing=TRUE), paired=TRUE) #significant p = 0.0076
#not paired
t.test(x=FR_sum, y = FR_win, paired= FALSE)#significant p= 0.0002
t.test(x=FR_sum, y = FR_win, paired= FALSE, var.equal = TRUE) #significant p=0.000196

#histograms and assumptions
hist(plant_order_FR$Poales, breaks = 10) #all observations, not normal
hist(FR_sum, breaks = 10)#summer, not normal
hist(FR_win, breaks = 10) #winter, not normal
hist(FR_sum - FR_win, breaks = 10) #difference , not normal

boxplot(plant_order_FR$Poales ~ plant_order_FR$season) #equalish variance 
var(FR_win) #0.045
var(FR_sum) #0.029

#GM
t.test(Poales ~ season, data= plant_order_GM)# not significant p= 0.1413

GM_sum<- plant_order_GM[plant_order_GM$season == "summer",]
GM_sum<-GM_sum$Poales
GM_win<- plant_order_GM[plant_order_GM$season == "winter",]
GM_win<-GM_win$Poales

#min variation
t.test(x=sort(GM_sum), y=sort(GM_win,decreasing=FALSE), paired=TRUE) #significant p= 0.000069
#max variation
t.test(x=sort(GM_sum), y=sort(GM_win,decreasing=TRUE), paired=TRUE)# not significant p= 0.2973
#not paired
t.test(x=GM_sum, y = GM_win, paired= FALSE) # not significant p= 0.1413
t.test(x=GM_sum, y = GM_win, paired= FALSE, var.equal = TRUE) #not significant p=0.1413

#histograms and assumptions
hist(plant_order_GM$Poales, breaks = 10) #all observations, not normal
hist(GM_sum, breaks = 10)#summer, not normal
hist(GM_win, breaks = 10) #winter, not normal
hist(GM_sum - GM_win, breaks = 10) #difference , not normal

boxplot(plant_order_GM$Poales ~ plant_order_GM$season) #equalish variance
var(GM_win) #0.076
var(GM_sum) #0.0804



#LBC
t.test(Poales ~ season, data= plant_order_LBC)#significant p=0.02412

LBC_sum<- plant_order_LBC[plant_order_LBC$season == "summer",]
LBC_sum<-LBC_sum$Poales
LBC_win<- plant_order_LBC[plant_order_LBC$season == "winter",]
LBC_win<-LBC_win$Poales

#min variation
t.test(x=sort(LBC_sum), y=sort(LBC_win,decreasing=FALSE), paired=TRUE) #significant p=0.00000008
#max variation
t.test(x=sort(LBC_sum), y=sort(LBC_win,decreasing=TRUE), paired=TRUE) #not significant p=0.1001
#not paired
t.test(x=LBC_sum, y = LBC_win, paired= FALSE) #significant p=0.02412
t.test(x=LBC_sum, y = LBC_win, paired= FALSE, var.equal = TRUE) #significant p=0.0239

#histograms and assumptions
hist(plant_order_LBC$Poales, breaks = 10) #all observations, not normal
hist(LBC_sum, breaks = 10)#summer, not normal
hist(LBC_win, breaks = 10) #winter, not normal


boxplot(plant_order_LBC$Poales ~ plant_order_LBC$season)#variance is not super equal, but not terrible, more varaince in winter
var(LBC_win) #0.119
var(LBC_sum) #0.086




#MP 
t.test(Poales ~ season, data= plant_order_MP)#not significant p=0.5294

MP_sum<- plant_order_MP[plant_order_MP$season == "summer",]
MP_sum<-MP_sum$Poales
MP_win<- plant_order_MP[plant_order_MP$season == "winter",]
MP_win<-MP_win$Poales

#min variation
t.test(x=sort(MP_sum), y=sort(MP_win,decreasing=FALSE), paired=TRUE) #borderline significant p=0.062
#max variation
t.test(x=sort(MP_sum), y=sort(MP_win,decreasing=TRUE), paired=TRUE)# not significant p=0.638
#not paired
t.test(x=MP_sum, y = MP_win, paired= FALSE) #not significant p= 0.5294
t.test(x=MP_sum, y = MP_win, paired= FALSE, var.equal = TRUE) #not significant p=0.5288

#histograms and assumptions
hist(plant_order_MP$Poales, breaks = 10) #all observations. not normal
hist(MP_sum, breaks = 10)#summer, not normal
hist(MP_win, breaks = 10) #winter, not normal
hist(MP_sum - MP_win, breaks = 10) #difference , normal-ish 

boxplot(plant_order_MP$Poales ~ plant_order_MP$season)# variance not super equal, more variance winter
var(MP_win)# 0.1323
var(MP_sum) #0.0679






#NH
t.test(Poales ~ season, data= plant_order_NH) #not significant p=0.1918

NH_sum<- plant_order_NH[plant_order_NH$season == "summer",]
NH_sum<-NH_sum$Poales
NH_win<- plant_order_NH[plant_order_NH$season == "winter",]
NH_win<-NH_win$Poales

#min variation
t.test(x=sort(NH_sum), y=sort(NH_win,decreasing=FALSE), paired=TRUE) #significant p=0.0022
#max variation
t.test(x=sort(NH_sum), y=sort(NH_win,decreasing=TRUE), paired=TRUE)#not significant p=0.3050
#not paired
t.test(x=NH_sum, y = NH_win, paired= FALSE)#not significant p=0.1918
t.test(x=NH_sum, y = NH_win, paired= FALSE, var.equal = TRUE) #not significant p=0.1909

#histograms and assumptions
hist(plant_order_NH$Poales, breaks = 10) #all observations, not normal
hist(NH_sum, breaks = 10)#summer, not normal
hist(NH_win, breaks = 10) #winter, not normal
hist(NH_sum - NH_win, breaks = 10) #difference, not normal 

boxplot(plant_order_NH$Poales ~ plant_order_NH$season) #variance not too equal, more variance winter, summer outliers
var(NH_win) #0.0819
var(NH_sum) #0.046




#ON
t.test(Poales ~ season, data= plant_order_ON)#significant p=0.0091

ON_sum<- plant_order_ON[plant_order_ON$season == "summer",]
ON_sum<-ON_sum$Poales
ON_win<- plant_order_ON[plant_order_ON$season == "winter",]
ON_win<-ON_win$Poales

#min variation
t.test(x=sort(ON_sum), y=sort(ON_win,decreasing=FALSE), paired=TRUE) #significant p=0.000003
#max variation
t.test(x=sort(ON_sum), y=sort(ON_win,decreasing=TRUE), paired=TRUE)#borderline significant p=0.05338
#not paired
t.test(x=ON_sum, y = ON_win, paired= FALSE)#significant p=0.0091
t.test(x=ON_sum, y = ON_win, paired= FALSE, var.equal = TRUE) #significant p=0.0089

#histograms and assumptions
hist(plant_order_ON$Poales, breaks = 10) #all observations, not normal
hist(ON_sum, breaks = 10)#summer, not normal
hist(ON_win, breaks = 10) #winter, not normal
hist(ON_sum - ON_win, breaks = 10) #difference , not normal

boxplot(plant_order_ON$Poales ~ plant_order_ON$season) #variance is not quite equal, more variance, summer, winter outliar 
var(ON_win) #0.0439
var(ON_sum) #0.0692

#PB
t.test(Poales ~ season, data=plant_order_PB) #significant p=0.0116

PB_sum<- plant_order_PB[plant_order_PB$season == "summer",]
PB_sum<-PB_sum$Poales
PB_win<- plant_order_PB[plant_order_PB$season == "winter",]
PB_win<-PB_win$Poales

#min variation
t.test(x=sort(PB_sum), y=sort(PB_win,decreasing=FALSE), paired=TRUE)#significant p= 0.0008 
#max variation
t.test(x=sort(PB_sum), y=sort(PB_win,decreasing=TRUE), paired=TRUE) #significant p = 0.03127
#not paired
t.test(x=PB_sum, y = PB_win, paired= FALSE) #significant p=0.0116
t.test(x=PB_sum, y = PB_win, paired= FALSE, var.equal = TRUE) #significant p=0.00846

#histograms and assumptions
hist(plant_order_PB$Poales, breaks = 10) #all observations, not normal
hist(PB_sum, breaks = 10)#summer, nor normal
hist(PB_win, breaks = 10) #winter, not normal
hist(PB_sum - PB_win, breaks = 10) #difference, closer to normal

boxplot(plant_order_PB$Poales ~ plant_order_PB$season)#not equal, more variance in summer, outliers in winter
var(PB_win) #0.0097
var(PB_sum) #0.102




#PN
t.test(Poales ~ season, data=plant_order_PN)# not significant p=0.6241

PN_sum<- plant_order_PN[plant_order_PN$season == "summer",]
PN_sum<-PN_sum$Poales
PN_win<- plant_order_PN[plant_order_PN$season == "winter",]
PN_win<-PN_win$Poales

#min variation
t.test(x=sort(PN_sum), y=sort(PN_win,decreasing=FALSE), paired=TRUE) #borderline significant p=0.0804
#max variation
t.test(x=sort(PN_sum), y=sort(PN_win,decreasing=TRUE), paired=TRUE)# not significant p=0.7212
#not paired
t.test(x=PN_sum, y = PN_win, paired= FALSE)# not significant p=0.6241
t.test(x=PN_sum, y = PN_win, paired= FALSE, var.equal = TRUE) #not significant p=0.6239

#histograms and assumptions
hist(plant_order_PN$Poales, breaks = 10) #all observations, not normal
hist(PN_sum, breaks = 10)#summer, not normal
hist(PN_win, breaks = 10) #winter, not normal
hist(PN_sum - PN_win, breaks = 10) #difference , not normal

boxplot(plant_order_PN$Poales ~ plant_order_PN$season)#equalish variance
var(PN_win)#0.046
var(PN_sum) #0.074




#RR
t.test(Poales ~ season, data=plant_order_RR)# not significant p=0.4158

RR_sum<- plant_order_RR[plant_order_RR$season == "summer",]
RR_sum<-RR_sum$Poales
RR_win<- plant_order_RR[plant_order_RR$season == "winter",]
RR_win<-RR_win$Poales

#min variation
t.test(x=sort(RR_sum), y=sort(RR_win,decreasing=FALSE), paired=TRUE) # significant p=0.00098
#max variation
t.test(x=sort(RR_sum), y=sort(RR_win,decreasing=TRUE), paired=TRUE)# not significant p=0.5658
#not paired
t.test(x=RR_sum, y = RR_win, paired= FALSE)# not significant p=0.4158
t.test(x=RR_sum, y = RR_win, paired= FALSE, var.equal = TRUE) #not significant p=0.4158

#histograms and assumptions
hist(plant_order_RR$Poales, breaks = 10) #all observations, normalish
hist(RR_sum, breaks = 10)#summer, not normal
hist(RR_win, breaks = 10) #winter, not terrible
hist(RR_sum - RR_win, breaks = 10) #difference #not normal

boxplot(plant_order_RR$Poales ~ plant_order_RR$season)#fairly equal variance
var(RR_win)#0.0688
var(RR_sum) #0.0768



#SC
t.test(Poales ~ season, data=plant_order_SC)# significant p=0.0003

SC_sum<- plant_order_SC[plant_order_SC$season == "summer",]
SC_sum<-SC_sum$Poales
SC_win<- plant_order_SC[plant_order_SC$season == "winter",]
SC_win<-SC_win$Poales

#min variation
t.test(x=sort(SC_sum), y=sort(SC_win,decreasing=FALSE), paired=TRUE) # significant p= 0.000001
#max variation
t.test(x=sort(SC_sum), y=sort(SC_win,decreasing=TRUE), paired=TRUE)# significant p= 0.0090
#not paired
t.test(x=SC_sum, y = SC_win, paired= FALSE)# significant p=0.0003
t.test(x=SC_sum, y = SC_win, paired= FALSE, var.equal = TRUE) #significant p=0.000299


#histograms and assumptions
hist(plant_order_SC$Poales, breaks = 10) #all observations, not normal
hist(SC_sum, breaks = 10)#summer, not normal
hist(SC_win, breaks = 10) #winter, not normal
hist(SC_sum - SC_win, breaks = 10) #difference , not normal

boxplot(plant_order_SC$Poales ~ plant_order_SC$season) #close variance, more variance winter, summer outliers
var(SC_win) #0.0535
var(SC_sum) #0.0366




#SS
t.test(Poales ~ season, data=plant_order_SS)#not significant p=0.6111

SS_sum<- plant_order_SS[plant_order_SS$season == "summer",]
SS_sum<-SS_sum$Poales
SS_win<- plant_order_SS[plant_order_SS$season == "winter",]
SS_win<-SS_win$Poales

#min variation
t.test(x=sort(SS_sum), y=sort(SS_win,decreasing=FALSE), paired=TRUE) #not significant p=0.1290
#max variation
t.test(x=sort(SS_sum), y=sort(SS_win,decreasing=TRUE), paired=TRUE)#not significant p=0.7122
#not paired
t.test(x=SS_sum, y = SS_win, paired= FALSE)#not significant p=0.6111
t.test(x=SS_sum, y = SS_win, paired= FALSE, var.equal = TRUE) #not significant p=0.611

#histograms and assumptions
hist(plant_order_SS$Poales, breaks = 10) #all observations, not terrible
hist(SS_sum, breaks = 10)#summer, not normal
hist(SS_win, breaks = 10) #winter, normalish
hist(SS_sum - SS_win, breaks = 10) #difference, not normal 

boxplot(plant_order_SS$Poales ~ plant_order_SS$season) #not super equal- winter has outliers, more variance summer 
var(SS_win) #0.0359
var(SS_sum) #0.0517

#SW
t.test(Poales ~ season, data=plant_order_SW)#borderline significant p=0.07915

SW_sum<- plant_order_SW[plant_order_SW$season == "summer",]
SW_sum<-SW_sum$Poales
SW_win<- plant_order_SW[plant_order_SW$season == "winter",]
SW_win<-SW_win$Poales

#min variation
t.test(x=sort(SW_sum), y=sort(SW_win,decreasing=FALSE), paired=TRUE) #significant p=0.0002
#max variation
t.test(x=sort(SW_sum), y=sort(SW_win,decreasing=TRUE), paired=TRUE)#not significant p=0.1995
#not paired
t.test(x=SW_sum, y = SW_win, paired= FALSE)#borderline significant p=0.07915
t.test(x=SW_sum, y = SW_win, paired= FALSE, var.equal = TRUE) #borderline significant p=0.07716

#histograms and assumptions
hist(plant_order_SW$Poales, breaks = 10) #all observations, not normal
hist(SW_sum, breaks = 10)#summer, not normal
hist(SW_win, breaks = 10) #winter, not normal
hist(SW_sum - SW_win, breaks = 10) #difference , not normal

boxplot(plant_order_SW$Poales ~ plant_order_SW$season)#not very equal, winter more variable 
var(SW_win) #0.088
var(SW_sum)#0.035



#TP
t.test(Poales ~ season, data=plant_order_TP)# not significant p=0.9896
TP_sum<- plant_order_TP[plant_order_TP$season == "summer",]
TP_sum<-TP_sum$Poales
TP_win<- plant_order_TP[plant_order_TP$season == "winter",]
TP_win<-TP_win$Poales

#min variation
t.test(x=sort(TP_sum), y=sort(TP_win,decreasing=FALSE), paired=TRUE) # not significant p=0.9519
#max variation
t.test(x=sort(TP_sum), y=sort(TP_win,decreasing=TRUE), paired=TRUE)# not significant p=0.9925
#not paired
t.test(x=TP_sum, y = TP_win, paired= FALSE)# not significant p=0.9896
t.test(x=TP_sum, y = TP_win, paired= FALSE, var.equal = TRUE) #not significant p=0.9896

#histograms and assumptions
hist(plant_order_TP$Poales, breaks = 10) #all observations, not normal
hist(TP_sum, breaks = 10)#summer, not normal
hist(TP_win, breaks = 10) #winter, not normal
hist(TP_sum - TP_win) #difference, not terrible... 

boxplot(plant_order_TP$Poales ~ plant_order_TP$season)#equalish, summer more variable 
var(TP_win) #0.0972
var(TP_sum) #0.113

```


## Bootstrapping

###individual HMAs

####independent (old name)

```{r}
#code from notes from STATS 5025

permutation.test<-function(obsdat, respvar, treatvar, numsims){
  Fperm=c()
  result=0
  for (i in 1:numsims){
    newtreat<-sample(treatvar, nrow(obsdat), replace =FALSE)
    mod<-lm(respvar ~ newtreat, data = obsdat)
    anovamod<-summary(aov(mod))
    Fperm[i] <- anovamod[[1]]$"F value"[1]
  }
result=sum(Fperm>= Fobs)
return(list(result, Fperm))
}


#AT
AT_anova<-lm(Poales~season, data= plant_order_AT)#original ANOVA
summary(aov(AT_anova)) #view anova summary
Fobs<-summary(aov(AT_anova)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
AT_perm1<- permutation.test(obsdat=plant_order_AT, respvar=plant_order_AT$Poales, treatvar=plant_order_AT$season, numsims=1000)

AT_perm1
AT_p_val<-AT_perm1[[1]]/1000
AT_p_val #p=0.861


#CN
#CN 
CN_anova<-lm(Poales~season, data= plant_order_CN)
summary(aov(CN_anova))
Fobs<-summary(aov(CN_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
CN_perm1<- permutation.test(obsdat=plant_order_CN, respvar=plant_order_CN$Poales, treatvar=plant_order_CN$season, numsims=1000)

CN_perm1
CN_p_val<-CN_perm1[[1]]/1000
CN_p_val #p=0.196

#FR

FR_anova<-lm(Poales~season, data= plant_order_FR)
summary(aov(FR_anova))
Fobs<-summary(aov(FR_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
FR_perm1<- permutation.test(obsdat=plant_order_FR, respvar=plant_order_FR$Poales, treatvar=plant_order_FR$season, numsims=1000)

FR_perm1
FR_p_val<-FR_perm1[[1]]/1000
FR_p_val #p=0.0000


#GM 
GM_anova<-lm(Poales~season, data= plant_order_GM)
summary(aov(GM_anova))
Fobs<-summary(aov(GM_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
GM_perm1<- permutation.test(obsdat=plant_order_GM, respvar=plant_order_GM$Poales, treatvar=plant_order_GM$season, numsims=1000)

GM_perm1
GM_p_val<-GM_perm1[[1]]/1000
GM_p_val #p=0.141

#LBC 
LBC_anova<-lm(Poales~season, data= plant_order_LBC)
summary(aov(LBC_anova))
Fobs<-summary(aov(LBC_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
LBC_perm1<- permutation.test(obsdat=plant_order_LBC, respvar=plant_order_LBC$Poales, treatvar=plant_order_LBC$season, numsims=1000)

LBC_perm1
LBC_p_val<-LBC_perm1[[1]]/1000
LBC_p_val #p=0.024


#MP 
MP_anova<-lm(Poales~season, data= plant_order_MP)
summary(aov(MP_anova))
Fobs<-summary(aov(MP_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
 
set.seed(223)  
MP_perm1<- permutation.test(obsdat=plant_order_MP, respvar=plant_order_MP$Poales, treatvar=plant_order_MP$season, numsims=1000)

MP_perm1
MP_p_val<-MP_perm1[[1]]/1000
MP_p_val #p=0.537


#NH 
NH_anova<-lm(Poales~season, data= plant_order_NH)
summary(aov(NH_anova))
Fobs<-summary(aov(NH_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223) 
NH_perm1<- permutation.test(obsdat=plant_order_NH, respvar=plant_order_NH$Poales, treatvar=plant_order_NH$season, numsims=1000)

NH_perm1
NH_p_val<-NH_perm1[[1]]/1000
NH_p_val #p=0.173

#ON 
ON_anova<-lm(Poales~season, data= plant_order_ON)
summary(aov(ON_anova))
Fobs<-summary(aov(ON_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)  
ON_perm1<- permutation.test(obsdat=plant_order_ON, respvar=plant_order_ON$Poales, treatvar=plant_order_ON$season, numsims=1000)

ON_perm1
ON_p_val<-ON_perm1[[1]]/1000
ON_p_val #p=0.006

#PB 
PB_anova<-lm(Poales~season, data= plant_order_PB)
summary(aov(PB_anova))
Fobs<-summary(aov(PB_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
PB_perm1<- permutation.test(obsdat=plant_order_PB, respvar=plant_order_PB$Poales, treatvar=plant_order_PB$season, numsims=1000)

PB_perm1
PB_p_val<-PB_perm1[[1]]/1000
PB_p_val #p=0.007

#PN 
PN_anova<-lm(Poales~season, data= plant_order_PN)
summary(aov(PN_anova))
Fobs<-summary(aov(PN_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
PN_perm1<- permutation.test(obsdat=plant_order_PN, respvar=plant_order_PN$Poales, treatvar=plant_order_PN$season, numsims=1000)

PN_perm1
PN_p_val<-PN_perm1[[1]]/1000
PN_p_val #p=0.630

#RR 
RR_anova<-lm(Poales~season, data= plant_order_RR)
summary(aov(RR_anova))
Fobs<-summary(aov(RR_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
RR_perm1<- permutation.test(obsdat=plant_order_RR, respvar=plant_order_RR$Poales, treatvar=plant_order_RR$season, numsims=1000)

RR_perm1
RR_p_val<-RR_perm1[[1]]/1000
RR_p_val #p=0.417

#SC 
SC_anova<-lm(Poales~season, data= plant_order_SC)
summary(aov(SC_anova))
Fobs<-summary(aov(SC_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SC_perm1<- permutation.test(obsdat=plant_order_SC, respvar=plant_order_SC$Poales, treatvar=plant_order_SC$season, numsims=1000)

SC_perm1
SC_p_val<-SC_perm1[[1]]/1000
SC_p_val #p=0.001


#SS 
SS_anova<-lm(Poales~season, data= plant_order_SS)
summary(aov(SS_anova))
Fobs<-summary(aov(SS_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SS_perm1<- permutation.test(obsdat=plant_order_SS, respvar=plant_order_SS$Poales, treatvar=plant_order_SS$season, numsims=1000)

SS_perm1
SS_p_val<-SS_perm1[[1]]/1000
SS_p_val #p=0.606

#SW 
SW_anova<-lm(Poales~season, data= plant_order_SW)
summary(aov(SW_anova))
Fobs<-summary(aov(SW_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SW_perm1<- permutation.test(obsdat=plant_order_SW, respvar=plant_order_SW$Poales, treatvar=plant_order_SW$season, numsims=1000)

SW_perm1
SW_p_val<-SW_perm1[[1]]/1000
SW_p_val #p=0.090

#TP 
TP_anova<-lm(Poales~season, data= plant_order_TP)
summary(aov(TP_anova))
Fobs<-summary(aov(TP_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  

set.seed(223)  
TP_perm1<- permutation.test(obsdat=plant_order_TP, respvar=plant_order_TP$Poales, treatvar=plant_order_TP$season, numsims=1000)

TP_perm1
TP_p_val<-TP_perm1[[1]]/1000
TP_p_val #p=0.988

```
#### independent bootstrapped (new name)

```{r}
#code from notes from STATS 5025

bootstrap_anova.test<-function(obsdat, respvar, treatvar, numsims){
  Fperm=c()
  result=0
  for (i in 1:numsims){
    newtreat<-sample(treatvar, nrow(obsdat), replace =FALSE)
    mod<-lm(respvar ~ newtreat, data = obsdat)
    anovamod<-summary(aov(mod))
    Fperm[i] <- anovamod[[1]]$"F value"[1]
  }
result=sum(Fperm>= Fobs)
return(list(result, Fperm))
}


#AT
AT_anova<-lm(Poales~season, data= plant_order_AT)#original ANOVA
summary(aov(AT_anova)) #view anova summary
Fobs<-summary(aov(AT_anova)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
AT_perm1<- bootstrap_anova.test(obsdat=plant_order_AT, respvar=plant_order_AT$Poales, treatvar=plant_order_AT$season, numsims=1000)

AT_perm1
AT_p_val<-AT_perm1[[1]]/1000
AT_p_val #p=0.861


#CN
#CN 
CN_anova<-lm(Poales~season, data= plant_order_CN)
summary(aov(CN_anova))
Fobs<-summary(aov(CN_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
CN_perm1<- bootstrap_anova.test(obsdat=plant_order_CN, respvar=plant_order_CN$Poales, treatvar=plant_order_CN$season, numsims=1000)

CN_perm1
CN_p_val<-CN_perm1[[1]]/1000
CN_p_val #p=0.196

#FR

FR_anova<-lm(Poales~season, data= plant_order_FR)
summary(aov(FR_anova))
Fobs<-summary(aov(FR_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
FR_perm1<- bootstrap_anova.test(obsdat=plant_order_FR, respvar=plant_order_FR$Poales, treatvar=plant_order_FR$season, numsims=1000)

FR_perm1
FR_p_val<-FR_perm1[[1]]/1000
FR_p_val #p=0.0000


#GM 
GM_anova<-lm(Poales~season, data= plant_order_GM)
summary(aov(GM_anova))
Fobs<-summary(aov(GM_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)   
GM_perm1<- bootstrap_anova.test(obsdat=plant_order_GM, respvar=plant_order_GM$Poales, treatvar=plant_order_GM$season, numsims=1000)

GM_perm1
GM_p_val<-GM_perm1[[1]]/1000
GM_p_val #p=0.141

#LBC 
LBC_anova<-lm(Poales~season, data= plant_order_LBC)
summary(aov(LBC_anova))
Fobs<-summary(aov(LBC_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
LBC_perm1<- bootstrap_anova.test(obsdat=plant_order_LBC, respvar=plant_order_LBC$Poales, treatvar=plant_order_LBC$season, numsims=1000)

LBC_perm1
LBC_p_val<-LBC_perm1[[1]]/1000
LBC_p_val #p=0.024


#MP 
MP_anova<-lm(Poales~season, data= plant_order_MP)
summary(aov(MP_anova))
Fobs<-summary(aov(MP_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
 
set.seed(223)  
MP_perm1<- bootstrap_anova.test(obsdat=plant_order_MP, respvar=plant_order_MP$Poales, treatvar=plant_order_MP$season, numsims=1000)

MP_perm1
MP_p_val<-MP_perm1[[1]]/1000
MP_p_val #p=0.537


#NH 
NH_anova<-lm(Poales~season, data= plant_order_NH)
summary(aov(NH_anova))
Fobs<-summary(aov(NH_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223) 
NH_perm1<- bootstrap_anova.test(obsdat=plant_order_NH, respvar=plant_order_NH$Poales, treatvar=plant_order_NH$season, numsims=1000)

NH_perm1
NH_p_val<-NH_perm1[[1]]/1000
NH_p_val #p=0.173

#ON 
ON_anova<-lm(Poales~season, data= plant_order_ON)
summary(aov(ON_anova))
Fobs<-summary(aov(ON_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs

set.seed(223)  
ON_perm1<- bootstrap_anova.test(obsdat=plant_order_ON, respvar=plant_order_ON$Poales, treatvar=plant_order_ON$season, numsims=1000)

ON_perm1
ON_p_val<-ON_perm1[[1]]/1000
ON_p_val #p=0.006

#PB 
PB_anova<-lm(Poales~season, data= plant_order_PB)
summary(aov(PB_anova))
Fobs<-summary(aov(PB_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
PB_perm1<- bootstrap_anova.test(obsdat=plant_order_PB, respvar=plant_order_PB$Poales, treatvar=plant_order_PB$season, numsims=1000)

PB_perm1
PB_p_val<-PB_perm1[[1]]/1000
PB_p_val #p=0.007

#PN 
PN_anova<-lm(Poales~season, data= plant_order_PN)
summary(aov(PN_anova))
Fobs<-summary(aov(PN_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
PN_perm1<- bootstrap_anova.test(obsdat=plant_order_PN, respvar=plant_order_PN$Poales, treatvar=plant_order_PN$season, numsims=1000)

PN_perm1
PN_p_val<-PN_perm1[[1]]/1000
PN_p_val #p=0.630

#RR 
RR_anova<-lm(Poales~season, data= plant_order_RR)
summary(aov(RR_anova))
Fobs<-summary(aov(RR_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
RR_perm1<- bootstrap_anova.test(obsdat=plant_order_RR, respvar=plant_order_RR$Poales, treatvar=plant_order_RR$season, numsims=1000)

RR_perm1
RR_p_val<-RR_perm1[[1]]/1000
RR_p_val #p=0.417

#SC 
SC_anova<-lm(Poales~season, data= plant_order_SC)
summary(aov(SC_anova))
Fobs<-summary(aov(SC_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SC_perm1<- bootstrap_anova.test(obsdat=plant_order_SC, respvar=plant_order_SC$Poales, treatvar=plant_order_SC$season, numsims=1000)

SC_perm1
SC_p_val<-SC_perm1[[1]]/1000
SC_p_val #p=0.001


#SS 
SS_anova<-lm(Poales~season, data= plant_order_SS)
summary(aov(SS_anova))
Fobs<-summary(aov(SS_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SS_perm1<- bootstrap_anova.test(obsdat=plant_order_SS, respvar=plant_order_SS$Poales, treatvar=plant_order_SS$season, numsims=1000)

SS_perm1
SS_p_val<-SS_perm1[[1]]/1000
SS_p_val #p=0.606

#SW 
SW_anova<-lm(Poales~season, data= plant_order_SW)
summary(aov(SW_anova))
Fobs<-summary(aov(SW_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  
set.seed(223)  
SW_perm1<- bootstrap_anova.test(obsdat=plant_order_SW, respvar=plant_order_SW$Poales, treatvar=plant_order_SW$season, numsims=1000)

SW_perm1
SW_p_val<-SW_perm1[[1]]/1000
SW_p_val #p=0.090

#TP 
TP_anova<-lm(Poales~season, data= plant_order_TP)
summary(aov(TP_anova))
Fobs<-summary(aov(TP_anova))
Fobs<-Fobs[[1]]$"F value"[1]
Fobs
  

set.seed(223)  
TP_perm1<- bootstrap_anova.test(obsdat=plant_order_TP, respvar=plant_order_TP$Poales, treatvar=plant_order_TP$season, numsims=1000)

TP_perm1
TP_p_val<-TP_perm1[[1]]/1000
TP_p_val #p=0.992
```
#### independent bootstrapped (boot.t.test)
```{r}
set.seed(223)
boot.t.test(x=AT_sum, y = AT_win, var.equal = TRUE) # p=0.851
set.seed(223)
boot.t.test(x=AT_sum, y = AT_win) # p=0.820

set.seed(223)
boot.t.test(x=CN_sum, y = CN_win, var.equal = TRUE) # p= 0.193
set.seed(223)
boot.t.test(x=CN_sum, y = CN_win)# p= 0.189

set.seed(223)
boot.t.test(x=FR_sum, y = FR_win, var.equal = TRUE) # p= 0.0001
set.seed(223)
boot.t.test(x=FR_sum, y = FR_win) # p=0.001

set.seed(223)
boot.t.test(x=GM_sum, y = GM_win, var.equal = TRUE) # p=0.1422
set.seed(223)
boot.t.test(x=GM_sum, y = GM_win) # p=0.1544

set.seed(223)
boot.t.test(x=LBC_sum, y = LBC_win, var.equal = TRUE) # p= 0.027
set.seed(223)
boot.t.test(x=LBC_sum, y = LBC_win) # p = 0.0258

set.seed(223)
boot.t.test(x=MP_sum, y = MP_win, var.equal = TRUE) # p= 0.5303
set.seed(223)
boot.t.test(x=MP_sum, y = MP_win) # p= 0.519

set.seed(223)
boot.t.test(x=NH_sum, y = NH_win, var.equal = TRUE) # p=0.1938
set.seed(223)
boot.t.test(x=NH_sum, y = NH_win) # p=0.1738

set.seed(223)
boot.t.test(x=ON_sum, y = ON_win, var.equal = TRUE) # p=0.0088
set.seed(223)
boot.t.test(x=ON_sum, y = ON_win) #p=0.017

set.seed(223)
boot.t.test(x=PB_sum, y = PB_win, var.equal = TRUE) # p= 0.0064
set.seed(223)
boot.t.test(x=PB_sum, y = PB_win) # p=0.0034

set.seed(223)
boot.t.test(x=PN_sum, y = PN_win, var.equal = TRUE) # p=0.6167
set.seed(223)
boot.t.test(x=PN_sum, y = PN_win) # p=0.6315

set.seed(223)
boot.t.test(x=RR_sum, y = RR_win, var.equal = TRUE) # p = 0.415
set.seed(223)
boot.t.test(x=RR_sum, y = RR_win) # p =0.4244

set.seed(223)
boot.t.test(x=SC_sum, y = SC_win, var.equal = TRUE) # p= 0.0001
set.seed(223)
boot.t.test(x=SC_sum, y = SC_win) # p=0.0032

set.seed(223)
boot.t.test(x=SS_sum, y = SS_win, var.equal = TRUE) #p= 0.6209
set.seed(223)
boot.t.test(x=SS_sum, y = SS_win) # p= 0.6183

set.seed(223)
boot.t.test(x=SW_sum, y = SW_win, var.equal = TRUE) # p=0.07501
set.seed(223)
boot.t.test(x=SW_sum, y = SW_win) # p=0.1022

set.seed(223)
boot.t.test(x=TP_sum, y = TP_win, var.equal = TRUE) # p=0.9763
set.seed(223)
boot.t.test(x=TP_sum, y = TP_win) # p=0.9981

```


#### paired


```{r}
#Use the MKinfer::boot.t.test 

#AT 

AT_sum #list summer graminoid compositions
AT_win # #list winter graminoid compositions

AT_min_dif<-sort(AT_sum, decreasing=FALSE)-sort(AT_win,decreasing=FALSE) #pair samples for the minimum variance between summer and winter sample differences- this represents the lowest possible p-value we could achieve if all samples were paired so that the lowest graminoid composition summer is paired with the lowest graminoid composition winter and the highest summer is paired with highest winter  
AT_min_dif

AT_max_dif<-sort(AT_sum, decreasing=FALSE)-sort(AT_win,decreasing=TRUE)
AT_max_dif #pair samples for the maximum variance between summer and winter sample differences- this represents the highest possible p-value we could achieve if all samples were paired so that the lowest graminoid composition summer is paired with the highest graminoid composition winter and the highest summer is paired with lowest winter  

set.seed(223)
boot.t.test(AT_min_dif) #p=0.6535

set.seed(223)
boot.t.test(AT_max_dif)#p=0.8749


#playing with the function to see if we get the same results with using the two vectors of summer and winter rather than just the difference and we get the same result 

#Min
set.seed(223)
boot.t.test(x=sort(AT_sum, decreasing=FALSE), y= sort(AT_win,decreasing=FALSE), paired=TRUE)

#Max
set.seed(223)
boot.t.test(x=sort(AT_sum, decreasing=FALSE), y= sort(AT_win,decreasing=TRUE), paired=TRUE)



#CN
CN_sum
CN_win

CN_min_dif<-sort(CN_sum, decreasing=FALSE)-sort(CN_win,decreasing=FALSE)
CN_min_dif

CN_max_dif<-sort(CN_sum, decreasing=FALSE)-sort(CN_win,decreasing=TRUE)
CN_max_dif


set.seed(223)
boot.t.test(CN_min_dif) #p=0.019

set.seed(223)
boot.t.test(CN_max_dif)#p=0.2926


#FR 

FR_sum
FR_win

FR_min_dif<-sort(FR_sum, decreasing=FALSE)-sort(FR_win,decreasing=FALSE)
FR_min_dif

FR_max_dif<-sort(FR_sum, decreasing=FALSE)-sort(FR_win,decreasing=TRUE)
FR_max_dif


set.seed(223)
boot.t.test(FR_min_dif) #p= < 0.001


set.seed(223)
boot.t.test(FR_max_dif)#p=0.0164


#GM 

GM_sum
GM_win

GM_min_dif<-sort(GM_sum, decreasing=FALSE)-sort(GM_win,decreasing=FALSE)
GM_min_dif

GM_max_dif<-sort(GM_sum, decreasing=FALSE)-sort(GM_win,decreasing=TRUE)
GM_max_dif


set.seed(223)
boot.t.test(GM_min_dif) #p= <0.001

set.seed(223)
boot.t.test(GM_max_dif)#p=0.3196



#LBC

LBC_sum
LBC_win

LBC_min_dif<-sort(LBC_sum, decreasing=FALSE)-sort(LBC_win,decreasing=FALSE)
LBC_min_dif

LBC_max_dif<-sort(LBC_sum, decreasing=FALSE)-sort(LBC_win,decreasing=TRUE)
LBC_max_dif


set.seed(223)
boot.t.test(LBC_min_dif) #p= <0.001

set.seed(223)
boot.t.test(LBC_max_dif)#p=0.1206



#MP

MP_sum
MP_win

MP_min_dif<-sort(MP_sum, decreasing=FALSE)-sort(MP_win,decreasing=FALSE)
MP_min_dif

MP_max_dif<-sort(MP_sum, decreasing=FALSE)-sort(MP_win,decreasing=TRUE)
MP_max_dif


set.seed(223)
boot.t.test(MP_min_dif) #p= 0.028

set.seed(223)
boot.t.test(MP_max_dif)#p=0.6173



#NH

NH_sum
NH_win

NH_min_dif<-sort(NH_sum, decreasing=FALSE)-sort(NH_win,decreasing=FALSE)
NH_min_dif

NH_max_dif<-sort(NH_sum, decreasing=FALSE)-sort(NH_win,decreasing=TRUE)
NH_max_dif


set.seed(223)
boot.t.test(NH_min_dif) #p= 0.p= <0.001

set.seed(223)
boot.t.test(NH_max_dif)#p=0.3088



#ON

ON_sum
ON_win

ON_min_dif<-sort(ON_sum, decreasing=FALSE)-sort(ON_win,decreasing=FALSE)
ON_min_dif

ON_max_dif<-sort(ON_sum, decreasing=FALSE)-sort(ON_win,decreasing=TRUE)
ON_max_dif


set.seed(223)
boot.t.test(ON_min_dif) #p= 0.0006

set.seed(223)
boot.t.test(ON_max_dif)#p=0.0836

#PB

PB_sum
PB_win

PB_min_dif<-sort(PB_sum, decreasing=FALSE)-sort(PB_win,decreasing=FALSE)
PB_min_dif

PB_max_dif<-sort(PB_sum, decreasing=FALSE)-sort(PB_win,decreasing=TRUE)
PB_max_dif


set.seed(223)
boot.t.test(PB_min_dif) #p= <0.0001

set.seed(223)
boot.t.test(PB_max_dif)#p=0.019


#PN

PN_sum
PN_win

PN_min_dif<-sort(PN_sum, decreasing=FALSE)-sort(PN_win,decreasing=FALSE)
PN_min_dif

PN_max_dif<-sort(PN_sum, decreasing=FALSE)-sort(PN_win,decreasing=TRUE)
PN_max_dif

set.seed(223)
boot.t.test(PN_min_dif) #p= 0.1316

set.seed(223)
boot.t.test(PN_max_dif)#p=0.7213


#RR

RR_sum
RR_win

RR_min_dif<-sort(RR_sum, decreasing=FALSE)-sort(RR_win,decreasing=FALSE)
RR_min_dif

RR_max_dif<-sort(RR_sum, decreasing=FALSE)-sort(RR_win,decreasing=TRUE)
RR_max_dif


set.seed(223)
boot.t.test(RR_min_dif) #p= 0.0056

set.seed(223)
boot.t.test(RR_max_dif)#p=0.5625



#SC

SC_sum
SC_win

SC_min_dif<-sort(SC_sum, decreasing=FALSE)-sort(SC_win,decreasing=FALSE)
SC_min_dif

SC_max_dif<-sort(SC_sum, decreasing=FALSE)-sort(SC_win,decreasing=TRUE)
SC_max_dif

set.seed(223)
boot.t.test(SC_min_dif) #p= <0.0001

set.seed(223)
boot.t.test(SC_max_dif)#p=0.0302



#SS

SS_sum
SS_win

SS_min_dif<-sort(SS_sum, decreasing=FALSE)-sort(SS_win,decreasing=FALSE)
SS_min_dif

SS_max_dif<-sort(SS_sum, decreasing=FALSE)-sort(SS_win,decreasing=TRUE)
SS_max_dif

set.seed(223)
boot.t.test(SS_min_dif) #p=0.1006

set.seed(223)
boot.t.test(SS_max_dif)#p=0.6905

#SW

SW_sum
SW_win

SW_min_dif<-sort(SW_sum, decreasing=FALSE)-sort(SW_win,decreasing=FALSE)
SW_min_dif


SW_max_dif<-sort(SW_sum, decreasing=FALSE)-sort(SW_win,decreasing=TRUE)
SW_max_dif

set.seed(223)
boot.t.test(SW_min_dif) #p=0.0014

set.seed(223)
boot.t.test(SW_max_dif)#p=0.2162


#TP

TP_sum
TP_win

TP_min_dif<-sort(TP_sum, decreasing=FALSE)-sort(TP_win,decreasing=FALSE)
TP_min_dif

TP_max_dif<-sort(TP_sum, decreasing=FALSE)-sort(TP_win,decreasing=TRUE)
TP_max_dif


set.seed(223)
boot.t.test(TP_min_dif) #p=0.9721

set.seed(223)
boot.t.test(TP_max_dif)#p=0.9811


```
###range

```{r}
#paired bootstraped t-test

#t-test paired...

range_sum # View values

range_win #view values

range_dif<- range_sum - range_win #make vesctor of difference of summer and winter

range_dif

set.seed(223)
boot.t.test(x=range_sum, y=range_win, paired=TRUE) #when they are paired we have a borderline significant p-value 0.050

set.seed(223)
boot.t.test(range_dif)

#independent non-paired bootstrap 

range_anova<-lm(mean_Poales~season, data= plant_order_by_HMA_season)#original ANOVA
summary(aov(range_anova)) #view anova summary
Fobs<-summary(aov(range_anova)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
range_perm1<- permutation.test(obsdat=plant_order_by_HMA_season, respvar=plant_order_by_HMA_season$mean_Poales, treatvar=plant_order_by_HMA_season$season, numsims=1000)

range_perm1
range_p_val<-range_perm1[[1]]/1000
range_p_val #p=0.221

```



# BCS seasonal comparison (to BCS)
##average BCS observations
```{r}

View(individual_BCS)#look at the file of individual observations
individual_BCS$body_condition_obs<-as.numeric(individual_BCS$body_condition_obs)#make BCS numeric



#summarize by HMA
BCS_by_HMA_season<-individual_BCS%>% #create a file that has summary statistics    
  group_by( study.area,area_season, season)%>%
  summarise(mean_BCS=mean(body_condition_obs), sd=sd(body_condition_obs), number= n())

View(BCS_by_HMA_season)#check values



#create a dataframe of just summer
BCS_by_HMA_summer<-subset(BCS_by_HMA_season, season == "Summer")
View(BCS_by_HMA_summer)#check the subset

#create a dataframe of just winter
BCS_by_HMA_winter<-subset(BCS_by_HMA_season, season == "Winter")
View(BCS_by_HMA_winter)#check the subset


#get some summary stats

sum(BCS_by_HMA_winter$number)#702

sum(BCS_by_HMA_summer$number)#707

#seasonal summary stats
#summer
mean(BCS_by_HMA_summer$number) # number of obs per HMA average 44.1875
sd(BCS_by_HMA_summer$number) # number of obs per HMA sd 21.98247

mean(subset(BCS_by_HMA_summer, study.area != "SB")$number) # excluding SB data number of obs per HMA average 45.5333 
sd(subset(BCS_by_HMA_summer, study.area != "SB")$number) # excluding SB data number of obs per HMA sd 22.06117

mean(BCS_by_HMA_summer$mean_BCS) # BCS HMA average 5.01898
mean(subset(BCS_by_HMA_summer, study.area != "SB")$mean_BCS) # excluding SB HMA average 5.014691 
sd(subset(BCS_by_HMA_summer, study.area != "SB")$mean_BCS) # excluding SB HMA sd 0.157855 


#winter

mean(BCS_by_HMA_winter$number) # number of obs per HMA average 46.8
sd(BCS_by_HMA_winter$number) # number of obs per HMA sd 18.04043

mean(BCS_by_HMA_winter$mean_BCS) # BCS HMA average 4.97807
sd(BCS_by_HMA_winter$mean_BCS) # BCS HMA sd 0.138982
```

## Visualizations of BCS
```{r}

#calculate SE for bar chart

BCS_by_HMA_season$number<-as.numeric(BCS_by_HMA_season$number)#make BCS number numeric

BCS_by_HMA_season$se<-BCS_by_HMA_season$sd/sqrt(BCS_by_HMA_season$number) #calculate SE


#barchart
body_condition_bar<- ggplot(BCS_by_HMA_season, aes(study.area, mean_BCS, fill=season)) + geom_bar(stat="identity", position=position_dodge()) + geom_errorbar(aes(ymin=mean_BCS-se, ymax=mean_BCS + se), width=.2, position=position_dodge(.9)) + scale_y_continuous(limits=c(0,6))

body_condition_bar + theme(text = element_text(size = 18, family= "Times New Roman")) + theme(legend.text = element_text(size = 18, color="black")) + theme(axis.text = element_text(size = 18, color= "black")) + theme(axis.title = element_text(size = 18, color= "black")) +scale_fill_manual(values=c("#db6d00","#006ddb"), name= "Season") + labs(title="Average Body Condition Score by HMA", x="HMA", y="Average Body Condition Score")+ theme(plot.title = element_text(hjust = 0.5)) 

#Histogram of body conditions

BCS_hist<- ggplot(individual_BCS, aes(x=body_condition_obs, fill=season))+geom_histogram(position=position_dodge(), binwidth = 1)


BCS_hist +theme_bw() + theme(text = element_text(size = 16, family= "Times New Roman")) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_fill_manual(values=c("#db6d00","#006ddb")) + labs(title="Body Condition Observations by Season", x="Body Condition Score", y="Count")+ theme(plot.title = element_text(hjust = 0.5)) 

ggsave("body_condition_season_hist.png", width=10, height=7, dpi=300)

#adding spaces 
BCS_hist<- ggplot(individual_BCS, aes(x=body_condition_obs, fill=season))+geom_histogram(position=position_dodge(), binwidth = 0.5)

BCS_hist +theme_bw() + theme(text = element_text(size = 16, family= "Times New Roman")) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_fill_manual(values=c("#db6d00","#006ddb")) + labs(title="Body Condition Observations by Season", x="Body Condition Score", y="Count")+ theme(plot.title = element_text(hjust = 0.5)) 

ggsave("body_condition_season_hist2.png", width=10, height=7, dpi=300)

##NOTE THIS CHUNK IS FOR A VISUAL THAT IS NOT WORKING YET
#stacked bar chart by HMA... work on this maybe- want a chart showing the proportion of each HMA with each observation numbers- how many 3, 4, 5 etc. 
BCS_stack_study_area<- ggplot(individual_BCS, aes(x=area_season, fill=body_condition_obs)) + geom_bar()

BCS_stack_study_area<- ggplot(individual_BCS, aes(x=area_season, y=body_condition_obs, fill=body_condition_obs)) + geom_bar(stat="identity", position="fill")

BCS_stack_study_area + theme(text = element_text(size = 12, family= "Times New Roman")) + theme(legend.text = element_text(size = 12, color="black")) + theme(axis.text = element_text(size = 12, color= "black")) + theme(axis.title = element_text(size = 12, color= "black"))  + labs(title="Body Condition Observations by Study Area", x="Body Condition Score", y="Count")+ theme(plot.title = element_text(hjust = 0.5)) + scale_fill_manual(values= c("#db6d00","#006ddb", "#ffff6d","#000000","#fdae61"))

##BCS to graminoid comparison 

```{r}

#Correlations of gramminoid composition to BCS on herd averages
########### summer ######################

#pull out summer values for BCS herd averages

BCS_by_HMA_summer<-subset(BCS_by_HMA_season, season == "Summer")
View(BCS_by_HMA_summer)#check the subset


graminoids_by_HMA_summer$mean_BCS<-as.numeric(BCS_by_HMA_summer$mean_BCS)#add BCS to graminoid composition
View(graminoids_by_HMA_summer)#check

#make one withoug SB to add to graminoids by_HMA_summer
BCS_by_HMA_summer2<-subset(BCS_by_HMA_summer, study.area != "SB")
View(BCS_by_HMA_summer2)


## with SB
ggplot(data=graminoids_by_HMA_summer, aes(x=mean_Poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Average BCS and Graminoids Composition Summer", x="% Graminoids diet", y="Herd BCS") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) 


ggsave("body_condition_graminoid_summer.png", width=10, height=7, dpi=300)

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$mean_BCS, method="spearman") #significant p=0.01893 rho = -0.578366

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$mean_BCS, method="pearson") #significant p=0.0353 cor = -0.5284

####without SB ##################

Fig5A<-ggplot(data=graminoids_by_HMA_summer2, aes(x=mean_Poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(x="Graminoids in Diet (%) Summer", y=" Summer Herd Average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=0.48, y= 4.6, label=" Spearman's rho = -0.554,  p= 0.035",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


ggsave("body_condition_graminoid_summer.png", width=10, height=7, dpi=300)

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$mean_BCS, method="spearman") #significant p=0.03491 rho = -0.5535

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$mean_BCS, method="pearson") #significant p=0.04607 cor = -0.5217


########### winter ######################

#pull out winter values for BCS herd averages

BCS_by_HMA_winter<-subset(BCS_by_HMA_season, season == "Winter")
View(BCS_by_HMA_winter)#check the subset

graminoids_by_HMA_winter$mean_BCS<-as.numeric(BCS_by_HMA_winter$mean_BCS)#add BCS to graminoid composition
View(graminoids_by_HMA_winter)#check


Fig5B<-ggplot(data=graminoids_by_HMA_winter, aes(x=mean_Poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(x="Graminoids in Diet (%) Winter", y="Winter Herd Average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=0.35, y= 4.6, label=" Spearman's rho = -0.118,  p= 0.676",color="black", size=8, family= "Times New Roman") + theme(text = element_text(family= "Times New Roman"))

ggsave("body_condition_graminoid_winter.png", width=10, height=7, dpi=300)

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$mean_BCS, method="spearman") # not significant p=0.6763 rho = -0.1179

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$mean_BCS, method="pearson") # not significant p=0.8561 cor = -0.0512

 

###### Summer grass and winter body condition #############

# put the summer poales value into and winter bcs dataframe

graminoids_by_HMA_summer2<-subset(graminoids_by_HMA_summer, area_season != "SB")# remove SB HMA because it was not collected in winter

View(graminoids_by_HMA_summer2)#check 

graminoids_by_HMA_winter$summer_Poales<-graminoids_by_HMA_summer2$mean_Poales #add summer graminoid composition to winter dataframe

View(graminoids_by_HMA_winter) #it worked

#corelations
ggplot(data=graminoids_by_HMA_winter, aes(x=summer_Poales, y=mean_BCS))+ geom_point() +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Average winter BCS and Graminoids consumed previous summer", x="% Poales diet previous summer", y="Herd BCS winter") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) 

Fig5C<-ggplot(data=graminoids_by_HMA_winter, aes(x=summer_Poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs( x="Graminoids in Diet Summer", y="Winter Herd Average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=0.48, y= 4.6, label=" Spearman's rho = -0.068,  p= 0.812",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

ggsave("body_condition_graminoid_sum_win.png", width=10, height=7, dpi=300)
  
cor.test(graminoids_by_HMA_winter$summer_Poales, graminoids_by_HMA_winter$mean_BCS, method="pearson") #not significant p=0.7211 cor= -0.10067 

cor.test(graminoids_by_HMA_winter$summer_Poales, graminoids_by_HMA_winter$mean_BCS, method="spearman") # not significant p=0.8124 rho = -0.067857

```
###Combine into Figure 5
```{r}
#use ggpubr
Figure5<-ggarrange(Fig5A, Fig5B,Fig5C, labels= c("A","B", "C"), ncol=2, nrow=2, common.legend=TRUE, legend="right", font.label = list(size = 24)) #combine Fig 5 A-C

Figure5 #view


ggsave("multipane_Fig5.png", width=18, height=12, dpi=300) #save

```

## seasonal BCS comparison
###individual HMAs(old name)
```{r}

#AT
AT_BCS<-subset(individual_BCS, study.area == "AT") #subset to only AT
AT_BCS %>% count(season) #get summer and winter observation counts
View(AT_BCS) #check data 

AT_anova1<-lm(body_condition_obs~season, data= AT_BCS)#original ANOVA
summary(aov(AT_anova1)) #view anova summary p=0.138
Fobs<-summary(aov(AT_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
AT_perm<- permutation.test(obsdat=AT_BCS, respvar=AT_BCS$body_condition_obs, treatvar=AT_BCS$season, numsims=1000)

AT_perm
AT_p_valBCS<-AT_perm[[1]]/1000
AT_p_valBCS #p=0.128

#CN
CN_BCS<-subset(individual_BCS, study.area == "CN") #subset to only CN
CN_BCS %>% count(season) #get summer and winter observation counts
View(CN_BCS) #check data 

CN_anova1<-lm(body_condition_obs~season, data= CN_BCS)#original ANOVA
summary(aov(CN_anova1)) #view anova summary p=0.71
Fobs<-summary(aov(CN_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
CN_perm<- permutation.test(obsdat=CN_BCS, respvar=CN_BCS$body_condition_obs, treatvar=CN_BCS$season, numsims=1000)

CN_perm
CN_p_valBCS<-CN_perm[[1]]/1000
CN_p_valBCS #p=0.982

#FR
FR_BCS<-subset(individual_BCS, study.area == "FR") #subset to only FR
FR_BCS %>% count(season) #get summer and winter observation counts
View(FR_BCS) #check data 

FR_anova1<-lm(body_condition_obs~season, data= FR_BCS)#original ANOVA
summary(aov(FR_anova1)) #view anova summary p=0.525
Fobs<-summary(aov(FR_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
FR_perm<- permutation.test(obsdat=FR_BCS, respvar=FR_BCS$body_condition_obs, treatvar=FR_BCS$season, numsims=1000)

FR_perm
FR_p_valBCS<-FR_perm[[1]]/1000
FR_p_valBCS #p=0.559

#GM
GM_BCS<-subset(individual_BCS, study.area == "GM") #subset to only GM
GM_BCS %>% count(season) #get summer and winter observation counts
View(GM_BCS) #check data 

GM_anova1<-lm(body_condition_obs~season, data= GM_BCS)#original ANOVA
summary(aov(GM_anova1)) #view anova summary p=0.614
Fobs<-summary(aov(GM_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
GM_perm<- permutation.test(obsdat=GM_BCS, respvar=GM_BCS$body_condition_obs, treatvar=GM_BCS$season, numsims=1000)

GM_perm
GM_p_valBCS<-GM_perm[[1]]/1000
GM_p_valBCS #p=0.713

#LBC
LBC_BCS<-subset(individual_BCS, study.area == "LBC") #subset to only LBC
LBC_BCS %>% count(season) #get summer and winter observation counts
View(LBC_BCS) #check data 

LBC_anova1<-lm(body_condition_obs~season, data= LBC_BCS)#original ANOVA
summary(aov(LBC_anova1)) #view anova summary p=0.399
Fobs<-summary(aov(LBC_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
LBC_perm<- permutation.test(obsdat=LBC_BCS, respvar=LBC_BCS$body_condition_obs, treatvar=LBC_BCS$season, numsims=1000)

LBC_perm
LBC_p_valBCS<-LBC_perm[[1]]/1000
LBC_p_valBCS #p=0.460

#MP
MP_BCS<-subset(individual_BCS, study.area == "MP") #subset to only MP
MP_BCS %>% count(season) #get summer and winter observation counts
View(MP_BCS) #check data 

MP_anova1<-lm(body_condition_obs~season, data= MP_BCS)#original ANOVA
summary(aov(MP_anova1)) #view anova summary p=0.834
Fobs<-summary(aov(MP_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
MP_perm<- permutation.test(obsdat=MP_BCS, respvar=MP_BCS$body_condition_obs, treatvar=MP_BCS$season, numsims=1000)

MP_perm
MP_p_valBCS<-MP_perm[[1]]/1000
MP_p_valBCS #p=0.852

#NH
NH_BCS<-subset(individual_BCS, study.area == "NH") #subset to only NH
NH_BCS %>% count(season) #get summer and winter observation counts
View(NH_BCS) #check data 

NH_anova1<-lm(body_condition_obs~season, data= NH_BCS)#original ANOVA
summary(aov(NH_anova1)) #view anova summary p=0.479
Fobs<-summary(aov(NH_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
NH_perm<- permutation.test(obsdat=NH_BCS, respvar=NH_BCS$body_condition_obs, treatvar=NH_BCS$season, numsims=1000)

NH_perm
NH_p_valBCS<-NH_perm[[1]]/1000
NH_p_valBCS #p=0.423

#ON
ON_BCS<-subset(individual_BCS, study.area == "ON") #subset to only ON
ON_BCS %>% count(season) #get summer and winter observation counts
View(ON_BCS) #check data 

ON_anova1<-lm(body_condition_obs~season, data= ON_BCS)#original ANOVA
summary(aov(ON_anova1)) #view anova summary p=0.230
Fobs<-summary(aov(ON_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
ON_perm<- permutation.test(obsdat=ON_BCS, respvar=ON_BCS$body_condition_obs, treatvar=ON_BCS$season, numsims=1000)

ON_perm
ON_p_valBCS<-ON_perm[[1]]/1000
ON_p_valBCS #p=0.232

#PB
PB_BCS<-subset(individual_BCS, study.area == "PB") #subset to only PB
PB_BCS %>% count(season) #get summer and winter observation counts
View(PB_BCS) #check data 

PB_anova1<-lm(body_condition_obs~season, data= PB_BCS)#original ANOVA
summary(aov(PB_anova1)) #view anova summary p=0.00886
Fobs<-summary(aov(PB_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
PB_perm<- permutation.test(obsdat=PB_BCS, respvar=PB_BCS$body_condition_obs, treatvar=PB_BCS$season, numsims=1000)

PB_perm
PB_p_valBCS<-PB_perm[[1]]/1000
PB_p_valBCS #p=0.016

#PN
PN_BCS<-subset(individual_BCS, study.area == "PN") #subset to only PN
PN_BCS %>% count(season) #get summer and winter observation counts
View(PN_BCS) #check data 

PN_anova1<-lm(body_condition_obs~season, data= PN_BCS)#original ANOVA
summary(aov(PN_anova1)) #view anova summary p=0.209
Fobs<-summary(aov(PN_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
PN_perm<- permutation.test(obsdat=PN_BCS, respvar=PN_BCS$body_condition_obs, treatvar=PN_BCS$season, numsims=1000)

PN_perm
PN_p_valBCS<-PN_perm[[1]]/1000
PN_p_valBCS #p=0.266

#RR
RR_BCS<-subset(individual_BCS, study.area == "RR") #subset to only RR
RR_BCS %>% count(season) #get summer and winter observation counts
View(RR_BCS) #check data 

RR_anova1<-lm(body_condition_obs~season, data= RR_BCS)#original ANOVA
summary(aov(RR_anova1)) #view anova summary p=0.676
Fobs<-summary(aov(RR_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
RR_perm<- permutation.test(obsdat=RR_BCS, respvar=RR_BCS$body_condition_obs, treatvar=RR_BCS$season, numsims=1000)

RR_perm
RR_p_valBCS<-RR_perm[[1]]/1000
RR_p_valBCS #p=0.742




#SB
SB_BCS<-subset(individual_BCS, study.area == "SB")#subset to only SB
SB_BCS %>% count(season)



#SC
SC_BCS<-subset(individual_BCS, study.area == "SC") #subset to only SC
SC_BCS %>% count(season) #get summer and winter observation counts
View(SC_BCS) #check data 

SC_anova1<-lm(body_condition_obs~season, data= SC_BCS)#original ANOVA
summary(aov(SC_anova1)) #view anova summary p=0.923
Fobs<-summary(aov(SC_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SC_perm<- permutation.test(obsdat=SC_BCS, respvar=SC_BCS$body_condition_obs, treatvar=SC_BCS$season, numsims=1000)

SC_perm
SC_p_valBCS<-SC_perm[[1]]/1000
SC_p_valBCS #p=0.784

#SS
SS_BCS<-subset(individual_BCS, study.area == "SS") #subset to only SS
SS_BCS %>% count(season) #get summer and winter observation counts
View(SS_BCS) #check data 

SS_anova1<-lm(body_condition_obs~season, data= SS_BCS)#original ANOVA
summary(aov(SS_anova1)) #view anova summary p=0.000131
Fobs<-summary(aov(SS_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SS_perm<- permutation.test(obsdat=SS_BCS, respvar=SS_BCS$body_condition_obs, treatvar=SS_BCS$season, numsims=1000)

SS_perm
SS_p_valBCS<-SS_perm[[1]]/1000
SS_p_valBCS #p=<0.001


#SW
SW_BCS<-subset(individual_BCS, study.area == "SW") #subset to only SW
SW_BCS %>% count(season) #get summer and winter observation counts
View(SW_BCS) #check data 

SW_anova1<-lm(body_condition_obs~season, data= SW_BCS)#original ANOVA
summary(aov(SW_anova1)) #view anova summary p=0.278
Fobs<-summary(aov(SW_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SW_perm<- permutation.test(obsdat=SW_BCS, respvar=SW_BCS$body_condition_obs, treatvar=SW_BCS$season, numsims=1000)

SW_perm
SW_p_valBCS<-SW_perm[[1]]/1000
SW_p_valBCS #p=0.329

#TP
TP_BCS<-subset(individual_BCS, study.area == "TP") #subset to only TP
TP_BCS %>% count(season) #get summer and winter observation counts
View(TP_BCS) #check data 

TP_anova1<-lm(body_condition_obs~season, data= TP_BCS)#original ANOVA
summary(aov(TP_anova1)) #view anova summary p=0.147
Fobs<-summary(aov(TP_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
TP_perm<- permutation.test(obsdat=TP_BCS, respvar=TP_BCS$body_condition_obs, treatvar=TP_BCS$season, numsims=1000)

TP_perm
TP_p_valBCS<-TP_perm[[1]]/1000
TP_p_valBCS #p=0.096


```
###individual HMA bootstrap (new name)

```{r}
#AT
AT_BCS<-subset(individual_BCS, study.area == "AT") #subset to only AT
AT_BCS %>% count(season) #get summer and winter observation counts
View(AT_BCS) #check data 

AT_anova1<-lm(body_condition_obs~season, data= AT_BCS)#original ANOVA
summary(aov(AT_anova1)) #view anova summary p=0.138
Fobs<-summary(aov(AT_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
AT_perm<- bootstrap_anova.test(obsdat=AT_BCS, respvar=AT_BCS$body_condition_obs, treatvar=AT_BCS$season, numsims=1000)

AT_perm
AT_p_valBCS<-AT_perm[[1]]/1000
AT_p_valBCS #p=0.128

#CN
CN_BCS<-subset(individual_BCS, study.area == "CN") #subset to only CN
CN_BCS %>% count(season) #get summer and winter observation counts
View(CN_BCS) #check data 

CN_anova1<-lm(body_condition_obs~season, data= CN_BCS)#original ANOVA
summary(aov(CN_anova1)) #view anova summary p=0.71
Fobs<-summary(aov(CN_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
CN_perm<- bootstrap_anova.test(obsdat=CN_BCS, respvar=CN_BCS$body_condition_obs, treatvar=CN_BCS$season, numsims=1000)

CN_perm
CN_p_valBCS<-CN_perm[[1]]/1000
CN_p_valBCS #p=0.982

#FR
FR_BCS<-subset(individual_BCS, study.area == "FR") #subset to only FR
FR_BCS %>% count(season) #get summer and winter observation counts
View(FR_BCS) #check data 

FR_anova1<-lm(body_condition_obs~season, data= FR_BCS)#original ANOVA
summary(aov(FR_anova1)) #view anova summary p=0.525
Fobs<-summary(aov(FR_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
FR_perm<- bootstrap_anova.test(obsdat=FR_BCS, respvar=FR_BCS$body_condition_obs, treatvar=FR_BCS$season, numsims=1000)

FR_perm
FR_p_valBCS<-FR_perm[[1]]/1000
FR_p_valBCS #p=0.559

#GM
GM_BCS<-subset(individual_BCS, study.area == "GM") #subset to only GM
GM_BCS %>% count(season) #get summer and winter observation counts
View(GM_BCS) #check data 

GM_anova1<-lm(body_condition_obs~season, data= GM_BCS)#original ANOVA
summary(aov(GM_anova1)) #view anova summary p=0.614
Fobs<-summary(aov(GM_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
GM_perm<- bootstrap_anova.test(obsdat=GM_BCS, respvar=GM_BCS$body_condition_obs, treatvar=GM_BCS$season, numsims=1000)

GM_perm
GM_p_valBCS<-GM_perm[[1]]/1000
GM_p_valBCS #p=0.713

#LBC
LBC_BCS<-subset(individual_BCS, study.area == "LBC") #subset to only LBC
LBC_BCS %>% count(season) #get summer and winter observation counts
View(LBC_BCS) #check data 

LBC_anova1<-lm(body_condition_obs~season, data= LBC_BCS)#original ANOVA
summary(aov(LBC_anova1)) #view anova summary p=0.399
Fobs<-summary(aov(LBC_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
LBC_perm<- bootstrap_anova.test(obsdat=LBC_BCS, respvar=LBC_BCS$body_condition_obs, treatvar=LBC_BCS$season, numsims=1000)

LBC_perm
LBC_p_valBCS<-LBC_perm[[1]]/1000
LBC_p_valBCS #p=0.460

#MP
MP_BCS<-subset(individual_BCS, study.area == "MP") #subset to only MP
MP_BCS %>% count(season) #get summer and winter observation counts
View(MP_BCS) #check data 

MP_anova1<-lm(body_condition_obs~season, data= MP_BCS)#original ANOVA
summary(aov(MP_anova1)) #view anova summary p=0.834
Fobs<-summary(aov(MP_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
MP_perm<- bootstrap_anova.test(obsdat=MP_BCS, respvar=MP_BCS$body_condition_obs, treatvar=MP_BCS$season, numsims=1000)

MP_perm
MP_p_valBCS<-MP_perm[[1]]/1000
MP_p_valBCS #p=0.852

#NH
NH_BCS<-subset(individual_BCS, study.area == "NH") #subset to only NH
NH_BCS %>% count(season) #get summer and winter observation counts
View(NH_BCS) #check data 

NH_anova1<-lm(body_condition_obs~season, data= NH_BCS)#original ANOVA
summary(aov(NH_anova1)) #view anova summary p=0.479
Fobs<-summary(aov(NH_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
NH_perm<- bootstrap_anova.test(obsdat=NH_BCS, respvar=NH_BCS$body_condition_obs, treatvar=NH_BCS$season, numsims=1000)

NH_perm
NH_p_valBCS<-NH_perm[[1]]/1000
NH_p_valBCS #p=0.423

#ON
ON_BCS<-subset(individual_BCS, study.area == "ON") #subset to only ON
ON_BCS %>% count(season) #get summer and winter observation counts
View(ON_BCS) #check data 

ON_anova1<-lm(body_condition_obs~season, data= ON_BCS)#original ANOVA
summary(aov(ON_anova1)) #view anova summary p=0.230
Fobs<-summary(aov(ON_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
ON_perm<- bootstrap_anova.test(obsdat=ON_BCS, respvar=ON_BCS$body_condition_obs, treatvar=ON_BCS$season, numsims=1000)

ON_perm
ON_p_valBCS<-ON_perm[[1]]/1000
ON_p_valBCS #p=0.232

#PB
PB_BCS<-subset(individual_BCS, study.area == "PB") #subset to only PB
PB_BCS %>% count(season) #get summer and winter observation counts
View(PB_BCS) #check data 

PB_anova1<-lm(body_condition_obs~season, data= PB_BCS)#original ANOVA
summary(aov(PB_anova1)) #view anova summary p=0.00886
Fobs<-summary(aov(PB_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
PB_perm<- bootstrap_anova.test(obsdat=PB_BCS, respvar=PB_BCS$body_condition_obs, treatvar=PB_BCS$season, numsims=1000)

PB_perm
PB_p_valBCS<-PB_perm[[1]]/1000
PB_p_valBCS #p=0.016

#PN
PN_BCS<-subset(individual_BCS, study.area == "PN") #subset to only PN
PN_BCS %>% count(season) #get summer and winter observation counts
View(PN_BCS) #check data 

PN_anova1<-lm(body_condition_obs~season, data= PN_BCS)#original ANOVA
summary(aov(PN_anova1)) #view anova summary p=0.209
Fobs<-summary(aov(PN_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
PN_perm<- bootstrap_anova.test(obsdat=PN_BCS, respvar=PN_BCS$body_condition_obs, treatvar=PN_BCS$season, numsims=1000)

PN_perm
PN_p_valBCS<-PN_perm[[1]]/1000
PN_p_valBCS #p=0.266

#RR
RR_BCS<-subset(individual_BCS, study.area == "RR") #subset to only RR
RR_BCS %>% count(season) #get summer and winter observation counts
View(RR_BCS) #check data 

RR_anova1<-lm(body_condition_obs~season, data= RR_BCS)#original ANOVA
summary(aov(RR_anova1)) #view anova summary p=0.676
Fobs<-summary(aov(RR_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
RR_perm<- bootstrap_anova.test(obsdat=RR_BCS, respvar=RR_BCS$body_condition_obs, treatvar=RR_BCS$season, numsims=1000)

RR_perm
RR_p_valBCS<-RR_perm[[1]]/1000
RR_p_valBCS #p=0.742




#SB
SB_BCS<-subset(individual_BCS, study.area == "SB")#subset to only SB
SB_BCS %>% count(season)



#SC
SC_BCS<-subset(individual_BCS, study.area == "SC") #subset to only SC
SC_BCS %>% count(season) #get summer and winter observation counts
View(SC_BCS) #check data 

SC_anova1<-lm(body_condition_obs~season, data= SC_BCS)#original ANOVA
summary(aov(SC_anova1)) #view anova summary p=0.923
Fobs<-summary(aov(SC_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SC_perm<- bootstrap_anova.test(obsdat=SC_BCS, respvar=SC_BCS$body_condition_obs, treatvar=SC_BCS$season, numsims=1000)

SC_perm
SC_p_valBCS<-SC_perm[[1]]/1000
SC_p_valBCS #p=0.784

#SS
SS_BCS<-subset(individual_BCS, study.area == "SS") #subset to only SS
SS_BCS %>% count(season) #get summer and winter observation counts
View(SS_BCS) #check data 

SS_anova1<-lm(body_condition_obs~season, data= SS_BCS)#original ANOVA
summary(aov(SS_anova1)) #view anova summary p=0.000131
Fobs<-summary(aov(SS_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SS_perm<- bootstrap_anova.test(obsdat=SS_BCS, respvar=SS_BCS$body_condition_obs, treatvar=SS_BCS$season, numsims=1000)

SS_perm
SS_p_valBCS<-SS_perm[[1]]/1000
SS_p_valBCS #p=<0.001


#SW
SW_BCS<-subset(individual_BCS, study.area == "SW") #subset to only SW
SW_BCS %>% count(season) #get summer and winter observation counts
View(SW_BCS) #check data 

SW_anova1<-lm(body_condition_obs~season, data= SW_BCS)#original ANOVA
summary(aov(SW_anova1)) #view anova summary p=0.278
Fobs<-summary(aov(SW_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
SW_perm<- bootstrap_anova.test(obsdat=SW_BCS, respvar=SW_BCS$body_condition_obs, treatvar=SW_BCS$season, numsims=1000)

SW_perm
SW_p_valBCS<-SW_perm[[1]]/1000
SW_p_valBCS #p=0.329

#TP
TP_BCS<-subset(individual_BCS, study.area == "TP") #subset to only TP
TP_BCS %>% count(season) #get summer and winter observation counts
View(TP_BCS) #check data 

TP_anova1<-lm(body_condition_obs~season, data= TP_BCS)#original ANOVA
summary(aov(TP_anova1)) #view anova summary p=0.147
Fobs<-summary(aov(TP_anova1)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
TP_perm<- bootstrap_anova.test(obsdat=TP_BCS, respvar=TP_BCS$body_condition_obs, treatvar=TP_BCS$season, numsims=1000)

TP_perm
TP_p_valBCS<-TP_perm[[1]]/1000
TP_p_valBCS #p=0.096


```
###individual HMA bootstrap (boot.t.test)
```{r}

#Use default bootstrapped t, var=T did not work in all HMAs I think because of unequal sample sizes?

#AT 
View(AT_BCS) #check data 

AT_BCS_sum<- AT_BCS[AT_BCS$season == "Summer",] #pull out summer observations
AT_BCS_sum<-AT_BCS_sum$body_condition_obs #pull data from just BCS column summer values


AT_BCS_win<- AT_BCS[AT_BCS$season == "Winter",] #pull out winter observations
AT_BCS_win<-AT_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=AT_BCS_sum, y = AT_BCS_win, var.equal = TRUE) # p=0.138

set.seed(223)
boot.t.test(x=AT_BCS_sum, y = AT_BCS_win) # p=0.1366

#CN 
View(CN_BCS) #check data 

CN_BCS_sum<- CN_BCS[CN_BCS$season == "Summer",] #pull out summer observations
CN_BCS_sum<-CN_BCS_sum$body_condition_obs #pull data from just BCS column summer values


CN_BCS_win<- CN_BCS[CN_BCS$season == "Winter",] #pull out winter observations
CN_BCS_win<-CN_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=CN_BCS_sum, y=CN_BCS_win, var.equal = T) # ERROR
set.seed(223)
boot.t.test(x=CN_BCS_sum, y = CN_BCS_win, var.equal = TRUE, R = 999) # ERROR
set.seed(223)
boot.t.test(x=CN_BCS_sum, y = CN_BCS_win, var.equal = TRUE, R = 99) # p=0.6263

set.seed(223)
boot.t.test(x=CN_BCS_sum, y=CN_BCS_win) # p= 0.7339


#FR 
View(FR_BCS) #check data 

FR_BCS_sum<- FR_BCS[FR_BCS$season == "Summer",] #pull out summer observations
FR_BCS_sum<-FR_BCS_sum$body_condition_obs #pull data from just BCS column summer values


FR_BCS_win<- FR_BCS[FR_BCS$season == "Winter",] #pull out winter observations
FR_BCS_win<-FR_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=FR_BCS_sum, y = FR_BCS_win, var.equal = TRUE) # ERROR

set.seed(223)
boot.t.test(x=FR_BCS_sum, y = FR_BCS_win, var.equal = TRUE, R = 999) # p=0.5105

set.seed(223)
boot.t.test(x=FR_BCS_sum, y = FR_BCS_win) # p= 0.4754

#GM 
View(GM_BCS) #check data 

GM_BCS_sum<- GM_BCS[GM_BCS$season == "Summer",] #pull out summer observations
GM_BCS_sum<-GM_BCS_sum$body_condition_obs #pull data from just BCS column summer values


GM_BCS_win<- GM_BCS[GM_BCS$season == "Winter",] #pull out winter observations
GM_BCS_win<-GM_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=GM_BCS_sum, y = GM_BCS_win, var.equal = TRUE) # p=0.6303

set.seed(223)
boot.t.test(x=GM_BCS_sum, y = GM_BCS_win) # p=0.6251

#LBC 
View(LBC_BCS) #check data 

LBC_BCS_sum<- LBC_BCS[LBC_BCS$season == "Summer",] #pull out summer observations
LBC_BCS_sum<-LBC_BCS_sum$body_condition_obs #pull data from just BCS column summer values


LBC_BCS_win<- LBC_BCS[LBC_BCS$season == "Winter",] #pull out winter observations
LBC_BCS_win<-LBC_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=LBC_BCS_sum, y = LBC_BCS_win, var.equal = TRUE) # p=0.4144
set.seed(223)
boot.t.test(x=LBC_BCS_sum, y = LBC_BCS_win) # p=0.251

#MP 
View(MP_BCS) #check data 

MP_BCS_sum<- MP_BCS[MP_BCS$season == "Summer",] #pull out summer observations
MP_BCS_sum<-MP_BCS_sum$body_condition_obs #pull data from just BCS column summer values


MP_BCS_win<- MP_BCS[MP_BCS$season == "Winter",] #pull out winter observations
MP_BCS_win<-MP_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=MP_BCS_sum, y = MP_BCS_win, var.equal = TRUE) # p=0.8397

set.seed(223)
boot.t.test(x=MP_BCS_sum, y = MP_BCS_win) # p=0.8381

#NH 
View(NH_BCS) #check data 

NH_BCS_sum<- NH_BCS[NH_BCS$season == "Summer",] #pull out summer observations
NH_BCS_sum<-NH_BCS_sum$body_condition_obs #pull data from just BCS column summer values


NH_BCS_win<- NH_BCS[NH_BCS$season == "Winter",] #pull out winter observations
NH_BCS_win<-NH_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=NH_BCS_sum, y = NH_BCS_win, var.equal = TRUE) # p=0.4868

set.seed(223)
boot.t.test(x=NH_BCS_sum, y = NH_BCS_win) # p=0.4608

#ON 
View(ON_BCS) #check data 

ON_BCS_sum<- ON_BCS[ON_BCS$season == "Summer",] #pull out summer observations
ON_BCS_sum<-ON_BCS_sum$body_condition_obs #pull data from just BCS column summer values


ON_BCS_win<- ON_BCS[ON_BCS$season == "Winter",] #pull out winter observations
ON_BCS_win<-ON_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=ON_BCS_sum, y = ON_BCS_win, var.equal = TRUE) # p=0.2374

set.seed(223)
boot.t.test(x=ON_BCS_sum, y = ON_BCS_win) # p=0.2636


#PB 
View(PB_BCS) #check data 

PB_BCS_sum<- PB_BCS[PB_BCS$season == "Summer",] #pull out summer observations
PB_BCS_sum<-PB_BCS_sum$body_condition_obs #pull data from just BCS column summer values


PB_BCS_win<- PB_BCS[PB_BCS$season == "Winter",] #pull out winter observations
PB_BCS_win<-PB_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=PB_BCS_sum, y = PB_BCS_win, var.equal = TRUE) # p=0.0072

set.seed(223)
boot.t.test(x=PB_BCS_sum, y = PB_BCS_win) # p=0.0026

#PN 
View(PN_BCS) #check data 

PN_BCS_sum<- PN_BCS[PN_BCS$season == "Summer",] #pull out summer observations
PN_BCS_sum<-PN_BCS_sum$body_condition_obs #pull data from just BCS column summer values


PN_BCS_win<- PN_BCS[PN_BCS$season == "Winter",] #pull out winter observations
PN_BCS_win<-PN_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=PN_BCS_sum, y = PN_BCS_win, var.equal = TRUE) # ERROR
set.seed(223)
boot.t.test(x=PN_BCS_sum, y = PN_BCS_win, var.equal = TRUE, R=999) # p=0.2242

set.seed(223)
boot.t.test(x=PN_BCS_sum, y = PN_BCS_win) # p=0.1762

#RR 
View(RR_BCS) #check data 

RR_BCS_sum<- RR_BCS[RR_BCS$season == "Summer",] #pull out summer observations
RR_BCS_sum<-RR_BCS_sum$body_condition_obs #pull data from just BCS column summer values


RR_BCS_win<- RR_BCS[RR_BCS$season == "Winter",] #pull out winter observations
RR_BCS_win<-RR_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=RR_BCS_sum, y = RR_BCS_win, var.equal = TRUE) # p=0.6995

set.seed(223)
boot.t.test(x=RR_BCS_sum, y = RR_BCS_win) # p=0.6977

#SC 
View(SC_BCS) #check data 

SC_BCS_sum<- SC_BCS[SC_BCS$season == "Summer",] #pull out summer observations
SC_BCS_sum<-SC_BCS_sum$body_condition_obs #pull data from just BCS column summer values


SC_BCS_win<- SC_BCS[SC_BCS$season == "Winter",] #pull out winter observations
SC_BCS_win<-SC_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=SC_BCS_sum, y = SC_BCS_win, var.equal = TRUE) # p=0.9081
set.seed(223)
boot.t.test(x=SC_BCS_sum, y = SC_BCS_win) # p=0.9317

#SS 
View(SS_BCS) #check data 

SS_BCS_sum<- SS_BCS[SS_BCS$season == "Summer",] #pull out summer observations
SS_BCS_sum<-SS_BCS_sum$body_condition_obs #pull data from just BCS column summer values


SS_BCS_win<- SS_BCS[SS_BCS$season == "Winter",] #pull out winter observations
SS_BCS_win<-SS_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=SS_BCS_sum, y = SS_BCS_win, var.equal = TRUE) # p=0.0001
set.seed(223)
boot.t.test(x=SS_BCS_sum, y = SS_BCS_win) # p=0.0001

#SW 
View(SW_BCS) #check data 

SW_BCS_sum<- SW_BCS[SW_BCS$season == "Summer",] #pull out summer observations
SW_BCS_sum<-SW_BCS_sum$body_condition_obs #pull data from just BCS column summer values


SW_BCS_win<- SW_BCS[SW_BCS$season == "Winter",] #pull out winter observations
SW_BCS_win<-SW_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=SW_BCS_sum, y = SW_BCS_win, var.equal = TRUE) # p=0.2778
set.seed(223)
boot.t.test(x=SW_BCS_sum, y = SW_BCS_win) # p=0.2732

#TP 
View(TP_BCS) #check data 

TP_BCS_sum<- TP_BCS[TP_BCS$season == "Summer",] #pull out summer observations
TP_BCS_sum<-TP_BCS_sum$body_condition_obs #pull data from just BCS column summer values


TP_BCS_win<- TP_BCS[TP_BCS$season == "Winter",] #pull out winter observations
TP_BCS_win<-TP_BCS_win$body_condition_obs #pull data from just BCS column winter values

set.seed(223)
boot.t.test(x=TP_BCS_sum, y = TP_BCS_win, var.equal = TRUE) # ERROR
set.seed(223)
boot.t.test(x=TP_BCS_sum, y = TP_BCS_win, var.equal = TRUE, R=999) # p=0.1381

set.seed(223)
boot.t.test(x=TP_BCS_sum, y = TP_BCS_win) # p=0.08001

```


###range
```{r}
#check the format of the dataframes we are using
View(BCS_by_HMA_season)
View(BCS_by_HMA_summer)

range_BCS_sum<-subset(BCS_by_HMA_summer, study.area != "SB")#remove SB because it is not paired
View(range_BCS_sum)#check
range_BCS_sum<-range_BCS_sum$mean_BCS #pull out only the BCS from data frame
range_BCS_sum #list of the 15 summer herd average BCS

range_BCS_win<-BCS_by_HMA_winter$mean_BCS #pull out only the BCS from winter data frame
range_BCS_win #list of the 15 winter herd average BCS


#paired

set.seed(223)
boot.t.test(x=range_BCS_sum, y=range_BCS_win, paired=TRUE) #when they are paired we have non significant p-value 0.205



#independent non-paired bootstrap 

range_BCS_anova<-lm(mean_BCS~season, data= BCS_by_HMA_season)#original ANOVA
summary(aov(range_anova)) #view anova summary p=0.213
Fobs<-summary(aov(range_anova)) #prep to pull F-statistic out of summary
Fobs<-Fobs[[1]]$"F value"[1] #pull F-statistic out of summary
Fobs

set.seed(223)  
range_perm_BCS<- permutation.test(obsdat=BCS_by_HMA_season, respvar=BCS_by_HMA_season$mean_BCS, treatvar=BCS_by_HMA_season$season, numsims=1000)

range_perm_BCS
rangeBCS_p_val<-range_perm_BCS[[1]]/1000
rangeBCS_p_val #p=0.217


```

#Graminoid diet to Herb Available Comparison
```{r}
View(HMA_RAP) #View to figure out which observtions correspond to study HMAs

HMA_RAP_subset<- HMA_RAP [c(122, 154, 146, 126, 118, 123, 147, 164, 85, 71, 42, 27, 25, 93, 135, 187),]#subset 16 HMAs of interest- this keeps them in alphabetical order to match the file we merge into  

View(HMA_RAP_subset) #check that I got the right ones
```

##Summer samples and herbaceous cover
```{r}
#add RAP columns to summer data for each HMA shape file extracted
graminoids_by_HMA_summer$RAP_AGF<-as.numeric(HMA_RAP_subset$annforbandgrasscov)#annual forbs and grass (herbaceous) cover

graminoids_by_HMA_summer$RAP_PGF<-as.numeric(HMA_RAP_subset$perforbandgrasscov)#perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_summer$RAP_total_GF<-as.numeric(graminoids_by_HMA_summer$RAP_AGF +graminoids_by_HMA_summer$RAP_PGF) #total annual plus perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_summer$RAP_shrub<-as.numeric(HMA_RAP_subset$shrub) #shrub cover

graminoids_by_HMA_summer$RAP_bare<-as.numeric(HMA_RAP_subset$bareground) #bare ground cover

graminoids_by_HMA_summer$RAP_annual_biomass<-as.numeric(HMA_RAP_subset$annforbandgrassbio)#annual forbs and grass (herbaceous) biomass

graminoids_by_HMA_summer$RAP_perennial_biomass<-as.numeric(HMA_RAP_subset$perforbandgrassbio)#perennial forbs and grass (herbaceous) biomass

graminoids_by_HMA_summer$RAP_herb_biomass<-as.numeric(HMA_RAP_subset$perforbandgrassbio + HMA_RAP_subset$annforbandgrassbio) #total forbs and grass (herbaceous) biomass

View(graminoids_by_HMA_summer)#check that all columns added in 


## Also do the ones that exclude SB

HMA_RAP_subset2<- subset(HMA_RAP_subset, OBJECTID != "27") #exclude SB from dataframe

View(HMA_RAP_subset2)#check

#add RAP columns to summer data for each HMA shape file extracted
graminoids_by_HMA_summer2$RAP_AGF<-as.numeric(HMA_RAP_subset2$annforbandgrasscov)#annual forbs and grass (herbaceous) cover

graminoids_by_HMA_summer2$RAP_PGF<-as.numeric(HMA_RAP_subset2$perforbandgrasscov)#perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_summer2$RAP_total_GF<-as.numeric(graminoids_by_HMA_summer2$RAP_AGF +graminoids_by_HMA_summer2$RAP_PGF) #total annual plus perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_summer2$RAP_shrub<-as.numeric(HMA_RAP_subset2$shrub) #shrub cover

graminoids_by_HMA_summer2$RAP_bare<-as.numeric(HMA_RAP_subset2$bareground) #bare ground cover

graminoids_by_HMA_summer2$RAP_annual_biomass<-as.numeric(HMA_RAP_subset2$annforbandgrassbio)#annual forbs and grass (herbaceous) biomass

graminoids_by_HMA_summer2$RAP_perennial_biomass<-as.numeric(HMA_RAP_subset2$perforbandgrassbio)#perennial forbs and grass (herbaceous) biomass

graminoids_by_HMA_summer2$RAP_herb_biomass<-as.numeric(HMA_RAP_subset2$perforbandgrassbio + HMA_RAP_subset2$annforbandgrassbio) #total forbs and grass (herbaceous) biomass

View(graminoids_by_HMA_summer2)#check that all columns added in 

graminoids_by_HMA_summer2[7, "RAP_total_GF"] <- 20.72217 #change the RAP herb cover for NH HMA to be weighted average for the HMA and WHT areas 

graminoids_by_HMA_summer2[7, "RAP_herb_biomass"] <- 267.20907 #change the total biomass for NH HMA to be weighted average for the HMA and WHT areas 


```


##summer herb cover and graminoid diet correlations
```{r}

############## full set of 16 HMA #######################

#add an ecoregion column to color points by 
graminoids_by_HMA_summer$epa_eco<- c("Wyoming Basin", "Central Basin", "Central Basin", "Wyoming Basin", "Colorado Plateau","Wyoming Basin", "Central Basin", "Central Basin", "Northern Basin", "Central Basin", "Mojave", "Northern Basin", "Snake River","Northern Basin", "Northern Basin","Northern Basin")

### Total grass forb cover
ggplot(data=graminoids_by_HMA_summer, aes(x=RAP_total_GF, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Herbaceous Cover and Graminoids consumed Summer", y="% Graminoids diet", x="% Herbaceous cover HMA") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$RAP_total_GF, method="spearman") #not significant p=0.1582 rho= -0.3706

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$RAP_total_GF, method="pearson") #not significant p=0.1640 cor= -.3654


### Total grass forb biomass
ggplot(data=graminoids_by_HMA_summer, aes(x=RAP_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Herbaceous Biomass and Graminoids consumed Summer", y="% Graminoids", x="% Herbaceous Biomass HMA") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$RAP_herb_biomass, method="spearman") #not significant p=0.1062 rho= -0.4206

cor.test(graminoids_by_HMA_summer$mean_Poales, graminoids_by_HMA_summer$RAP_herb_biomass, method="pearson") #not significant p=0.2052 cor= -.3346



############ Without sand basin ###############

graminoids_by_HMA_summer2$epa_eco<- c("Wyoming Basin", "Central Basin", "Central Basin", "Wyoming Basin", "Colorado Plateau","Wyoming Basin", "Central Basin", "Central Basin", "Northern Basin", "Central Basin", "Mojave",  "Snake River","Northern Basin", "Northern Basin","Northern Basin")

### Total grass forb cover
Fig3A<-ggplot(data=graminoids_by_HMA_summer2, aes(x=RAP_total_GF, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Summer", x="HMA Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=29, y=0.28, label=" Spearman's rho = -0.379,  p= 0.165",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$RAP_total_GF, method="spearman") #not significant p=0.1649 rho= -0.3786

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$RAP_total_GF, method="pearson") #not significant p=0.3241 cor= -.2734

ggsave("herb_cover_graminoid_summer.png", width=10, height=7, dpi=300)



### Total grass forb biomass
Fig3B<-ggplot(data=graminoids_by_HMA_summer2, aes(x=RAP_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Summer", x="HMA Herbaceous Biomass (lbs/acre)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=390, y=0.25, label=" Spearman's rho = -0.439,  p= 0.103",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$RAP_herb_biomass, method="spearman") #not significant p=0.1032 rho= -0.439

cor.test(graminoids_by_HMA_summer2$mean_Poales, graminoids_by_HMA_summer2$RAP_herb_biomass, method="pearson") #not significant p=0.4265 cor= -.2220

ggsave("herb_bio_graminoid_summer.png", width=10, height=7, dpi=300)

```
##Winter samples and herbaceous cover
```{r}


#add RAP columns to winter data for each HMA shape file extracted
graminoids_by_HMA_winter$RAP_AGF<-as.numeric(HMA_RAP_subset2$annforbandgrasscov)#annual forbs and grass (herbaceous) cover

graminoids_by_HMA_winter$RAP_PGF<-as.numeric(HMA_RAP_subset2$perforbandgrasscov)#perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_winter$RAP_total_GF<-as.numeric(graminoids_by_HMA_winter$RAP_AGF +graminoids_by_HMA_winter$RAP_PGF) #total annual plus perennial forbs and grass (herbaceous) cover

graminoids_by_HMA_winter$RAP_shrub<-as.numeric(HMA_RAP_subset2$shrub) #shrub cover

graminoids_by_HMA_winter$RAP_bare<-as.numeric(HMA_RAP_subset2$bareground) #bare ground cover

graminoids_by_HMA_winter$RAP_annual_biomass<-as.numeric(HMA_RAP_subset2$annforbandgrassbio)#annual forbs and grass (herbaceous) biomass

graminoids_by_HMA_winter$RAP_perennial_biomass<-as.numeric(HMA_RAP_subset2$perforbandgrassbio)#perennial forbs and grass (herbaceous) biomass

graminoids_by_HMA_winter$RAP_herb_biomass<-as.numeric(HMA_RAP_subset2$perforbandgrassbio + HMA_RAP_subset2$annforbandgrassbio) #total forbs and grass (herbaceous) biomass

View(graminoids_by_HMA_winter)#check that all columns added in 

graminoids_by_HMA_winter[7, "RAP_total_GF"] <- 20.72217 #change the RAP herb cover for NH HMA to be weighted average for the HMA and WHT areas 

graminoids_by_HMA_winter[7, "RAP_herb_biomass"] <- 267.20907 #change the total biomass for NH HMA to be weighted average for the HMA and WHT areas 

```


##winter herb cover and graminoid diet correlations
```{r}
graminoids_by_HMA_winter$epa_eco<- c("Wyoming Basin", "Central Basin", "Central Basin", "Wyoming Basin", "Colorado Plateau","Wyoming Basin", "Central Basin", "Central Basin", "Northern Basin", "Central Basin", "Mojave", "Snake River","Northern Basin", "Northern Basin","Northern Basin")

### Total grass forb cover
Fig3E<-ggplot(data=graminoids_by_HMA_winter, aes(x=RAP_total_GF, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Winter", x="HMA Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=29, y=0.05, label=" Spearman's rho = -0.279,  p= 0.314",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


ggsave("herb_cover_graminoid_winter.png", width=10, height=7, dpi=300)

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$RAP_total_GF, method="spearman") #not significant p=0.3138 rho= -0.2785

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$RAP_total_GF, method="pearson") #not significant p=0.3539 cor= -.2576



### Total grass forb biomass
Fig3F<- ggplot(data=graminoids_by_HMA_winter, aes(x=RAP_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Winter", x="HMA Herbaceous Biomass (lbs/acre)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=395, y=0.05, label=" Spearman's rho = -0.371,  p= 0.174",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

ggsave("herb_bio_graminoid_winter.png", width=10, height=7, dpi=300)

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$RAP_herb_biomass, method="spearman") #not significant p=0.1735 rho= -0.3714

cor.test(graminoids_by_HMA_winter$mean_Poales, graminoids_by_HMA_winter$RAP_herb_biomass, method="pearson") #not significant p=0.2912 cor= -.29187
```
## Individual buffer data
###prep files

```{r}

# Use the file Buffered each fecal location point by 13.5km diameter to represent 36 hours of movement with a 9km a day average

View(plant_family_meta_wild)#check what we are joining


metadata_with_RAP<- subset(metadata_with_RAP, domestication == "wild") #remove domestic horses so we have the same number of samples to join for our analyses


metadata_with_RAP<-subset(metadata_with_RAP, sample != "WMP11")#Exclude WMP11 as above
View(metadata_with_RAP) #check 

metadata_RAP_family<- left_join(plant_family_meta_wild, metadata_with_RAP, by="sample") #join the RAP data to individual sample data

View(metadata_RAP_family)#check 

metadata_RAP_family$RAP_total_GF<-metadata_RAP_family$perforbandgrasscov + metadata_RAP_family$annforbandgrasscov #make column of total herbaceous cover (annual + perennial grass and forb cover)

metadata_RAP_family$RAP_herb_biomass<-metadata_RAP_family$perforbandgrassbio + metadata_RAP_family$annforbandgrassbio #make column of total herbaceous biomass (annual + perennial grass and forb biomass)

#get the average percent cover for each HMA by averaging the buffers 

RAP_and_poales_by_HMA<-metadata_RAP_family%>% 
  group_by(study.area.x, area_season.x, season.x)%>%
  summarise(mean_Poales=mean(Poales), mean_herb_cover=mean(RAP_total_GF), mean_herb_biomass=mean(RAP_herb_biomass))
            
View(RAP_and_poales_by_HMA) #check 

RAP_and_poales_by_HMA_summer<-subset(RAP_and_poales_by_HMA, season.x == "summer")#subset to summer only

#add an ecoregion column to color points by later
RAP_and_poales_by_HMA_summer$epa_eco<- c("Wyoming Basin", "Central Basin", "Central Basin", "Wyoming Basin", "Colorado Plateau","Wyoming Basin", "Central Basin", "Central Basin", "Northern Basin", "Central Basin", "Mojave", "Northern Basin", "Snake River","Northern Basin", "Northern Basin","Northern Basin")

View(RAP_and_poales_by_HMA_summer) #check 

RAP_and_poales_by_HMA_winter<-subset(RAP_and_poales_by_HMA, season.x == "winter")#subset to winter only 

#add an ecoregion column to color points by later
RAP_and_poales_by_HMA_winter$epa_eco<- c("Wyoming Basin", "Central Basin", "Central Basin", "Wyoming Basin", "Colorado Plateau","Wyoming Basin", "Central Basin", "Central Basin", "Northern Basin", "Central Basin", "Mojave",  "Snake River","Northern Basin", "Northern Basin","Northern Basin")

View(RAP_and_poales_by_HMA_winter) #check 

```





###Summer
```{r}

########### With SB ######################

#Cover 

ggplot(data=RAP_and_poales_by_HMA_summer, aes(x=mean_herb_cover, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of Herd Average Buffered Herbaceous Cover and Graminoids consumed Summer", y="% Graminoids diet", x="% Average Herd Buffered Herbaceous Cover") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_cover, RAP_and_poales_by_HMA_summer$mean_Poales, method="spearman")#borderline significant p=0.05085 rho = -0.500


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_cover, RAP_and_poales_by_HMA_summer$mean_Poales, method="pearson")#borderline significant p=0.0778 cor = -0.4534

#Biomass

ggplot(data=RAP_and_poales_by_HMA_summer, aes(x=mean_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of Herd Average Buffered Herbaceous Biomass and Graminoids consumed Summer", y="% Graminoids diet", x="% Average Herd Buffered Herbaceous Biomass") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))



cor.test(RAP_and_poales_by_HMA_summer$mean_herb_biomass, RAP_and_poales_by_HMA_summer$mean_Poales, method="spearman")#borderline significant p=0.05558 rho = -0.49117


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_biomass, RAP_and_poales_by_HMA_summer$mean_Poales, method="pearson")#borderline significant p=0.07053 cor = -0.4635


########### Without SB ######################

RAP_and_poales_by_HMA_summer2<-subset(RAP_and_poales_by_HMA_summer, study.area.x != "SB") #exclude SB from dataframe

View(RAP_and_poales_by_HMA_summer2)#check


#Cover 

Fig3C<-ggplot(data=RAP_and_poales_by_HMA_summer2, aes(x=mean_herb_cover, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Summer", x="Herd Average Buffered Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=32.25, y=0.24, label=" Spearman's rho = -0.475,  p= 0.076",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


ggsave("herb_cover_graminoid_summer_buffer.png", width=10, height=7, dpi=300)

cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_cover, RAP_and_poales_by_HMA_summer2$mean_Poales, method="spearman")#borderline significant p=0.07587 rho = -0.475



cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_cover, RAP_and_poales_by_HMA_summer2$mean_Poales, method="pearson")#not significant p=0.1863 cor = -0.3609

#Biomass

Fig3D<-ggplot(data=RAP_and_poales_by_HMA_summer2, aes(x=mean_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Summer", x= "Herd Average Buffered Herbaceous Biomass (lbs/acre)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=425, y=0.22, label=" Spearman's rho = -0.479,  p= 0.073",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


ggsave("herb_bio_graminoid_summer_buffer.png", width=10, height=7, dpi=300)

cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_biomass, RAP_and_poales_by_HMA_summer2$mean_Poales, method="spearman")#borderline significant p=0.07347 rho = -0.47857


cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_biomass, RAP_and_poales_by_HMA_summer2$mean_Poales, method="pearson")#borderline significant p=0.1672 cor = -0.376

```


###Winter
```{r}
#Cover 

Fig3G<-ggplot(data=RAP_and_poales_by_HMA_winter, aes(x=mean_herb_cover, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Winter", x="Herd Average Buffered Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=32.5, y=0.05, label=" Spearman's rho = -0.268,  p= 0.333",color="black", size=8,  family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

ggsave("herb_cover_graminoid_winter_buffer.png", width=10, height=7, dpi=300)

cor.test(RAP_and_poales_by_HMA_winter$mean_herb_cover, RAP_and_poales_by_HMA_winter$mean_Poales, method="spearman")#not significant p=0.3334 rho = -0.2679


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_cover, RAP_and_poales_by_HMA_winter$mean_Poales, method="pearson")#not significant p=0.7178 rho = -0.10189

#Biomass

Fig3H<-ggplot(data=RAP_and_poales_by_HMA_winter, aes(x=mean_herb_biomass, y=mean_Poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Graminoids in Diet (%) Winter", x="Herd Average Buffered Herbaceous Biomass (lbs/acre)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=425, y=0.05, label=" Spearman's rho = -0.286,  p= 0.301",color="black", size = 8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


ggsave("herb_bio_graminoid_winter_buffer.png", width=10, height=7, dpi=300)

cor.test(RAP_and_poales_by_HMA_winter$mean_herb_biomass, RAP_and_poales_by_HMA_winter$mean_Poales, method="spearman")#not significant p=0.3011 rho = -0.2857


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_biomass, RAP_and_poales_by_HMA_winter$mean_Poales, method="pearson")#not significant p=0.3827 cor = -0.24309
```
### Combine Fig 3

```{r}

#use ggpubr
Figure3<-ggarrange(Fig3A, Fig3E , Fig3B, Fig3F, Fig3C, Fig3G, Fig3D, Fig3H, labels= c("A","E", "B", "F", "C", "G", "D", "H"), ncol=2, nrow=4, common.legend=TRUE, legend="right", font.label = list(size = 24)) #Combine Fig3A-H

Figure3 #View


ggsave("multipane_Fig3new.png", width=15, height=20, dpi=300)#save
```


### Individual Horses for Visual only 
```{r}
View(metadata_RAP_family)


ggplot(data=metadata_RAP_family, aes(x=RAP_total_GF, y=Poales))+ geom_point(aes(color=study.area.y, shape=season.y), size=5) + theme_bw()+labs(title = "Individual Horse Buffered Herbaceous Cover and Graminoids in Diet by HMA and Season", y= "Graminoids in Diet (%)", x=" Movement Buffer Herbaceous Cover") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) + scale_color_manual(values=c("#db6d00","#006ddb", "#e94c1f", "#000000","#fdae61", "#009292", "#24ff24","#ff6db6","#b6dbff",
 "#490092","#b66dff",
 "#920000", "#6db6ff","#004949",  "#924900","#ffb6db", "#FFFF00",  "#f7f7f7" ))+ theme(text = element_text(family= "Times New Roman"))

ggsave("individual_herb__graminoid.png", width=10, height=7, dpi=300)


#correlations
cor.test(metadata_RAP_family$RAP_total_GF, metadata_RAP_family$Poales, method="spearman") #significant p= 1.85 e-5 rho = -0.1972

cor.test(metadata_RAP_family$RAP_total_GF, metadata_RAP_family$Poales, method="pearson") #significant p=0.00068 cor = -0.1568

```
#### logit individuals 
```{r}
#fix values that are > 1 to 1, replace 1 to 0.99999 and 0 to 0.000001
metadata_RAP_family2<-metadata_RAP_family
View(metadata_RAP_family2)
metadata_RAP_family2[265, 74] = 1.00000
metadata_RAP_family2[59, 74] = 1.00000
metadata_RAP_family2[34, 74] = 1.00000

#These 2 lines of code makes the scales for the 0 and 1 values get weird... can run without/ with these lines and results are similar.  If run without the logit function adjusts for the 0 and 1
metadata_RAP_family2$Poales[metadata_RAP_family2$Poales==1]<-0.99999
metadata_RAP_family2$Poales[metadata_RAP_family2$Poales==0]<-0.0000001

metadata_RAP_family2$logit_poales<-logit(p=metadata_RAP_family2$Poales, percents=FALSE)#transform
View(metadata_RAP_family2)#check



ggplot(data=metadata_RAP_family2, aes(x=RAP_total_GF, y=logit_poales))+ geom_point(aes(color=study.area.y, shape=season.y), size=5) + theme_bw()+labs(title = "Individual Horse Buffered Herbaceous Cover and Graminoids in Diet logit by HMA and Season", y= "Logit Graminoids in Diet", x=" Movement Buffer Herbaceous Cover") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) + scale_color_manual(values=c("#db6d00","#006ddb", "#e94c1f", "#000000","#fdae61", "#009292", "#24ff24","#ff6db6","#b6dbff",
 "#490092","#b66dff",
 "#920000", "#6db6ff","#004949",  "#924900","#ffb6db", "#FFFF00",  "#f7f7f7" ))+ theme(text = element_text(family= "Times New Roman"))

#correlations
cor.test(metadata_RAP_family2$RAP_total_GF, metadata_RAP_family2$logit_poales, method="spearman") #significant p= 1.84 e-5 rho = -0.1972

cor.test(metadata_RAP_family2$RAP_total_GF, metadata_RAP_family2$logit_poales, method="pearson") #significant p=0.00019 cor = -0.17179


# Set to run without SB
metadata_RAP_family2<-metadata_RAP_family
metadata_RAP_family2<-subset(metadata_RAP_family2, study.area.y != "SB")

#fix values that are > 1 to 1, replace 1 to 0.99999 and 0 to 0.000001

View(metadata_RAP_family2)
metadata_RAP_family2[250, 74] = 1.00000
metadata_RAP_family2[59, 74] = 1.00000
metadata_RAP_family2[34, 74] = 1.00000

#These 2 lines of code makes the scales for the 0 and 1 values get weird... can run without/ with these lines and results are similar.  If run without the logit function adjusts for the 0 and 1 
metadata_RAP_family2$Poales[metadata_RAP_family2$Poales==1]<-0.99999
metadata_RAP_family2$Poales[metadata_RAP_family2$Poales==0]<-0.0000001

metadata_RAP_family2$logit_poales<-logit(p=metadata_RAP_family2$Poales, percents=FALSE)#transform
View(metadata_RAP_family2)#check



ggplot(data=metadata_RAP_family2, aes(x=RAP_total_GF, y=logit_poales))+ geom_point(aes(color=study.area.y, shape=season.y), size=5) + theme_bw()+labs(title = "Individual Horse Buffered Herbaceous Cover and Graminoids in Diet logit by HMA and Season", y= "Logit Graminoids in Diet", x=" Movement Buffer Herbaceous Cover") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) + scale_color_manual(values=c("#db6d00","#006ddb", "#e94c1f", "#000000","#fdae61", "#009292", "#24ff24","#ff6db6","#b6dbff",
 "#490092","#b66dff",
 "#920000", "#6db6ff","#004949",  "#924900","#ffb6db", "#FFFF00",  "#f7f7f7" ))+ theme(text = element_text(family= "Times New Roman")) + geom_smooth(method='lm')

#correlations
cor.test(metadata_RAP_family2$RAP_total_GF, metadata_RAP_family2$logit_poales, method="spearman") #significant p= 0.000127 rho = -0.17959

cor.test(metadata_RAP_family2$RAP_total_GF, metadata_RAP_family2$logit_poales, method="pearson") #significant p=0.002605 cor = -0.1416



```



#Logit transformed comparisons
## gramioids and BCS
```{r}

#Correlations of gramminoid composition to BCS on herd averages Logit transformed

########### summer ######################

#check dataframes to use
View(BCS_by_HMA_summer)#check the subset

View(graminoids_by_HMA_summer)#check

View(BCS_by_HMA_summer2)#check

View(graminoids_by_HMA_summer2)#check

#logit transform graminoid proportions

graminoids_by_HMA_summer$logit_poales<-logit(p=graminoids_by_HMA_summer$mean_Poales, percents=FALSE)#transform
View(graminoids_by_HMA_summer)#check

graminoids_by_HMA_summer2$logit_poales<-logit(p=graminoids_by_HMA_summer2$mean_Poales, percents=FALSE)#transform
View(graminoids_by_HMA_summer2)#check

## with SB
ggplot(data=graminoids_by_HMA_summer, aes(x=logit_poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Average BCS and Graminoids Composition Summer", x=" Logit graminoids diet", y="Herd BCS") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) 


cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$mean_BCS, method="spearman") #significant p=0.01893 rho = -0.578366

cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$mean_BCS, method="pearson") #significant p=0.02909 cor = -0.54482

####without SB ################## THIS IS ONE USED IN PAPER

ggplot(data=graminoids_by_HMA_summer2, aes(x=logit_poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(x="Logit graminoids in diet summer", y=" Summer herd average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=0.48, y= 4.6, label=" Spearman's rho = -0.554,  p= 0.035",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$mean_BCS, method="spearman") #significant p=0.03491 rho = -0.5535

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$mean_BCS, method="pearson") #significant p=0.0383 cor = -0.5385586


########### winter ######################

#check dataframes

View(BCS_by_HMA_winter)#check  
View(graminoids_by_HMA_winter)#check

#logit transform graminoid proportions

graminoids_by_HMA_winter$logit_poales<-logit(p=graminoids_by_HMA_winter$mean_Poales, percents=FALSE)#transform
View(graminoids_by_HMA_winter)#check


ggplot(data=graminoids_by_HMA_winter, aes(x=logit_poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(x="Logit graminoids in diet winter", y="Winter herd average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=0.35, y= 4.6, label=" Spearman's rho = -0.118,  p= 0.676",color="black", size=8, family= "Times New Roman") + theme(text = element_text(family= "Times New Roman"))


cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$mean_BCS, method="spearman") # not significant p=0.6763 rho = -0.1179

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$mean_BCS, method="pearson") # not significant p=0.9322 cor = -0.02405

 

###### Summer grass and winter body condition #############

# check that the summer poales value is in winter dataframe

View(graminoids_by_HMA_winter) #yes

#logit transform graminoid proportions

graminoids_by_HMA_winter$summer_logit_poales<-logit(p=graminoids_by_HMA_winter$summer_Poales, percents=FALSE)#transform
View(graminoids_by_HMA_winter)#check

#corelations

ggplot(data=graminoids_by_HMA_winter, aes(x=summer_logit_poales, y=mean_BCS, color=epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco, size=5)) +geom_smooth(method=lm, color="black")+ theme_bw()+labs( x="Logit graminoids in diet summer", y="Winter herd average BCS") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=0.48, y= 4.6, label=" Spearman's rho = -0.068,  p= 0.812",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

  
cor.test(graminoids_by_HMA_winter$summer_logit_poales, graminoids_by_HMA_winter$mean_BCS, method="pearson") #not significant p=0.6513 cor= -0.1272 

cor.test(graminoids_by_HMA_winter$summer_logit_poales, graminoids_by_HMA_winter$mean_BCS, method="spearman") # not significant p=0.8124 rho = -0.067857
```
## gramioids and cover/biomass
###Summer
```{r}
################## HMA average correlations ###################################

############## full set of 16 HMA #######################

View(graminoids_by_HMA_summer)#make sure it is the dataframe that I want

### Total grass forb cover
ggplot(data=graminoids_by_HMA_summer, aes(x=RAP_total_GF, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Herbaceous Cover and Graminoids consumed Summer", y="Logit Graminoids diet", x="% Herbaceous cover HMA") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))

cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$RAP_total_GF, method="spearman") #not significant p=0.1582 rho= -0.3706

cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$RAP_total_GF, method="pearson") #not significant p=0.1471 cor= -.37952


### Total grass forb biomass
ggplot(data=graminoids_by_HMA_summer, aes(x=RAP_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of HMA Herbaceous Biomass and Graminoids consumed Summer", y="Logit Graminoids", x="% Herbaceous Biomass HMA") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))

cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$RAP_herb_biomass, method="spearman") #not significant p=0.1062 rho= -0.4206

cor.test(graminoids_by_HMA_summer$logit_poales, graminoids_by_HMA_summer$RAP_herb_biomass, method="pearson") #not significant p=0.1839 cor= -.34994



############ Without sand basin ###############
View(graminoids_by_HMA_summer2) #check it is correct dataframe

### Total grass forb cover
ggplot(data=graminoids_by_HMA_summer2, aes(x=RAP_total_GF, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x="HMA Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=29, y=0.28, label=" Spearman's rho = -0.379,  p= 0.165",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$RAP_total_GF, method="spearman") #not significant p=0.1649 rho= -0.3786

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$RAP_total_GF, method="pearson") #not significant p=0.2753 cor= -.3012


### Total grass forb biomass
ggplot(data=graminoids_by_HMA_summer2, aes(x=RAP_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x="HMA Herbaceous Biomass (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=390, y=0.25, label=" Spearman's rho = -0.439,  p= 0.103",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$RAP_herb_biomass, method="spearman") #not significant p=0.1032 rho= -0.439

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$RAP_herb_biomass, method="pearson") #not significant p=0.3636 cor= -.2526



############### Buffered HMA correlations ##################################

########### With SB ######################

View(RAP_and_poales_by_HMA_summer) #check dataframe

RAP_and_poales_by_HMA_summer$logit_poales<-logit(p=RAP_and_poales_by_HMA_summer$mean_Poales, percents=FALSE)#transform
View(RAP_and_poales_by_HMA_summer)#check

#Cover 

ggplot(data=RAP_and_poales_by_HMA_summer, aes(x=mean_herb_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of Herd Average Buffered Herbaceous Cover and Logit Graminoids consumed Summer", y="Logit Graminoids diet", x="% Average Herd Buffered Herbaceous Cover") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_cover, RAP_and_poales_by_HMA_summer$logit_poales, method="spearman")#borderline significant p=0.05085 rho = -0.500


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_cover, RAP_and_poales_by_HMA_summer$logit_poales, method="pearson")#borderline significant p=0.06945 cor = -0.4651

#Biomass

ggplot(data=RAP_and_poales_by_HMA_summer, aes(x=mean_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(title = "Correlation of Herd Average Buffered Herbaceous Biomass and Logit Graminoids consumed Summer", y="Logit Graminoids diet", x="% Average Herd Buffered Herbaceous Biomass") + theme(text = element_text(size = 16)) + theme(plot.title = element_text(hjust = 0.5, size = 16)) + theme(legend.text = element_text(size = 16, color="black")) + theme(axis.text = element_text(size = 16, color= "black")) + theme(axis.title = element_text(size = 16, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))



cor.test(RAP_and_poales_by_HMA_summer$mean_herb_biomass, RAP_and_poales_by_HMA_summer$logit_poales, method="spearman")#borderline significant p=0.05558 rho = -0.49117


cor.test(RAP_and_poales_by_HMA_summer$mean_herb_biomass, RAP_and_poales_by_HMA_summer$logit_poales, method="pearson")#borderline significant p=0.06174 cor = -0.476985


########### Without SB ###################### 


View(RAP_and_poales_by_HMA_summer2)#check dataframe

RAP_and_poales_by_HMA_summer2$logit_poales<-logit(p=RAP_and_poales_by_HMA_summer2$mean_Poales, percents=FALSE)#transform
View(RAP_and_poales_by_HMA_summer2)#check


#Cover 

ggplot(data=RAP_and_poales_by_HMA_summer2, aes(x=mean_herb_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x="Herd Average Buffered Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=32.25, y=0.24, label=" Spearman's rho = -0.475,  p= 0.076",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))



cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_cover, RAP_and_poales_by_HMA_summer2$logit_poales, method="spearman")#borderline significant p=0.07587 rho = -0.475



cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_cover, RAP_and_poales_by_HMA_summer2$logit_poales, method="pearson")#not significant p=0.1581 cor = -0.383589

#Biomass

ggplot(data=RAP_and_poales_by_HMA_summer2, aes(x=mean_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x= "Herd Average Buffered Herbaceous Biomass") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + annotate(geom="text", x=425, y=0.22, label=" Spearman's rho = -0.479,  p= 0.073",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))



cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_biomass, RAP_and_poales_by_HMA_summer2$logit_poales, method="spearman")#borderline significant p=0.07347 rho = -0.47857


cor.test(RAP_and_poales_by_HMA_summer2$mean_herb_biomass, RAP_and_poales_by_HMA_summer2$logit_poales, method="pearson")#not significant p=0.1369 cor = -0.4025

```
###Winter
```{r}
############# HMA averages correlations #########################

View(graminoids_by_HMA_winter)

### Total grass forb cover
ggplot(data=graminoids_by_HMA_winter, aes(x=RAP_total_GF, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="HMA Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=29, y=0.05, label=" Spearman's rho = -0.279,  p= 0.314",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$RAP_total_GF, method="spearman") #not significant p=0.3138 rho= -0.2785

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$RAP_total_GF, method="pearson") #not significant p=0.4058 cor= -.23178



### Total grass forb biomass
ggplot(data=graminoids_by_HMA_winter, aes(x=RAP_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="HMA Herbaceous Biomass") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=395, y=0.05, label=" Spearman's rho = -0.371,  p= 0.174",color="black", size=8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$RAP_herb_biomass, method="spearman") #not significant p=0.1735 rho= -0.3714

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$RAP_herb_biomass, method="pearson") #not significant p=0.3462 cor= -.2616

############### Buffered HMA correlations ##################################

View(RAP_and_poales_by_HMA_winter) #check which dataframe
RAP_and_poales_by_HMA_winter$logit_poales<-logit(p=RAP_and_poales_by_HMA_winter$mean_Poales, percents=FALSE)#transform
View(RAP_and_poales_by_HMA_winter)#check

#Cover 

ggplot(data=RAP_and_poales_by_HMA_winter, aes(x=mean_herb_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="% Herd Average Buffered Herbaceous Cover (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=32.5, y=0.05, label=" Spearman's rho = -0.268,  p= 0.333",color="black", size=8,  family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_cover, RAP_and_poales_by_HMA_winter$logit_poales, method="spearman")#not significant p=0.3334 rho = -0.2679


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_cover, RAP_and_poales_by_HMA_winter$logit_poales, method="pearson")#not significant p=0.820 rho = -0.0643

#Biomass

ggplot(data=RAP_and_poales_by_HMA_winter, aes(x=mean_herb_biomass, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="% Herd Average Buffered Herbaceous Biomass (%)") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ annotate(geom="text", x=425, y=0.05, label=" Spearman's rho = -0.286,  p= 0.301",color="black", size = 8, family= "Times New Roman")+ theme(text = element_text(family= "Times New Roman"))


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_biomass, RAP_and_poales_by_HMA_winter$logit_poales, method="spearman")#not significant p=0.3011 rho = -0.2857


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_biomass, RAP_and_poales_by_HMA_winter$logit_poales, method="pearson")#not significant p=0.471 cor = -0.202
```
### Transformed Cover 
```{r}
######### SUMMER ################ 
################## HMA average correlations ###################################
#(no SB)

View(graminoids_by_HMA_summer2) #check it is correct dataframe
graminoids_by_HMA_summer2$logit_cover<- (graminoids_by_HMA_summer2$RAP_total_GF)/100 #turn cover % into proportion
graminoids_by_HMA_summer2$logit_cover<-logit(p=graminoids_by_HMA_summer2$logit_cover, percents=FALSE)#transform
View(graminoids_by_HMA_summer2)#check


### Total grass forb cover
ggplot(data=graminoids_by_HMA_summer2, aes(x=logit_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x="Logit Herbaceous Cover") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + theme(text = element_text(family= "Times New Roman"))

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$logit_cover, method="spearman") #not significant p=0.1649 rho= -0.3786

cor.test(graminoids_by_HMA_summer2$logit_poales, graminoids_by_HMA_summer2$logit_cover, method="pearson") #not significant p=0.2151 cor= -.339944


############### Buffered HMA correlations ##################################
########### Without SB ###################### 


View(RAP_and_poales_by_HMA_summer2)#check dataframe

RAP_and_poales_by_HMA_summer2$logit_cover<- (RAP_and_poales_by_HMA_summer2$mean_herb_cover)/100 #turn cover % into proportion

RAP_and_poales_by_HMA_summer2$logit_cover<-logit(p=RAP_and_poales_by_HMA_summer2$logit_cover, percents=FALSE)#transform
View(RAP_and_poales_by_HMA_summer2)#check


#Cover 

ggplot(data=RAP_and_poales_by_HMA_summer2, aes(x=logit_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Summer", x="Logit Herd Average Herbaceous Cover") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18)) + theme(text = element_text(family= "Times New Roman"))



cor.test(RAP_and_poales_by_HMA_summer2$logit_cover, RAP_and_poales_by_HMA_summer2$logit_poales, method="spearman")#borderline significant p=0.07587 rho = -0.475



cor.test(RAP_and_poales_by_HMA_summer2$logit_cover, RAP_and_poales_by_HMA_summer2$logit_poales, method="pearson")#not significant p=0.1202 cor = -0.4188


######## Winter ######################

############# HMA averages correlations #########################

View(graminoids_by_HMA_winter) #check which dataframe I am using 

graminoids_by_HMA_winter$logit_cover<- (graminoids_by_HMA_winter$RAP_total_GF)/100 #turn cover % into proportion
graminoids_by_HMA_winter$logit_cover<-logit(p=graminoids_by_HMA_winter$logit_cover, percents=FALSE)#transform
View(graminoids_by_HMA_winter)#check

### Total grass forb cover
ggplot(data=graminoids_by_HMA_winter, aes(x=logit_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="Logit Herbaceous Cover") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ theme(text = element_text(family= "Times New Roman"))


cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$logit_cover, method="spearman") #not significant p=0.3138 rho= -0.2785

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$logit_cover, method="pearson") #not significant p=0.411 cor= -.2293


############### Buffered HMA correlations ##################################

View(RAP_and_poales_by_HMA_winter) #check which dataframe
RAP_and_poales_by_HMA_winter$logit_cover<- (RAP_and_poales_by_HMA_winter$mean_herb_cover)/100 #turn cover % into proportion
RAP_and_poales_by_HMA_winter$logit_cover<-logit(p=RAP_and_poales_by_HMA_winter$logit_cover, percents=FALSE)#transform
View(RAP_and_poales_by_HMA_winter)#check

#Cover 

ggplot(data=RAP_and_poales_by_HMA_winter, aes(x=logit_cover, y=logit_poales, color= epa_eco))+ geom_point(aes(shape=epa_eco, color=epa_eco), size=5) +geom_smooth(method=lm, color="black")+ theme_bw()+labs(y="Logit Graminoids in Diet Winter", x="Logit Herd Buffered Herbaceous Cover") + theme(text = element_text(size = 20)) + theme(plot.title = element_text(hjust = 0.5, size = 20)) + theme(legend.text = element_text(size = 20, color="black")) + theme(axis.text = element_text(size = 20, color= "black")) + theme(axis.title = element_text(size = 20, color= "black")) +scale_color_manual(name="EPA III Ecoregion", values=c("#000000", "#009E73", "#E69F00", "#56B4E9", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) + scale_shape_manual(values=c(1,7,15,16,17,18))+ theme(text = element_text(family= "Times New Roman"))


cor.test(RAP_and_poales_by_HMA_winter$logit_cover, RAP_and_poales_by_HMA_winter$logit_poales, method="spearman")#not significant p=0.3334 rho = -0.2679


cor.test(RAP_and_poales_by_HMA_winter$logit_cover, RAP_and_poales_by_HMA_winter$logit_poales, method="pearson")#not significant p=0.761 rho = -0.0857

```
#Logit transformed linear regressions
## Response transformed Graminoids, predictor cover/biomass
```{r}
##Summer (no SB) 

#### Cover

View(graminoids_by_HMA_summer2) #check dataframe
mod1 = lm(formula= logit_poales ~ RAP_total_GF, data= graminoids_by_HMA_summer2)
summary(mod1)

#### Biomass

mod2 = lm(formula= logit_poales ~ RAP_herb_biomass, data= graminoids_by_HMA_summer2)
summary(mod2)

####Herd Buffered cover
View(RAP_and_poales_by_HMA_summer2) #check dataframe


mod3 = lm(formula= logit_poales ~ mean_herb_cover, data= RAP_and_poales_by_HMA_summer2)
summary(mod3)

#### Herd buffered biomass


mod4 = lm(formula= logit_poales ~ mean_herb_biomass, data= RAP_and_poales_by_HMA_summer2)
summary(mod4)

# Winter 

#### Cover

View(graminoids_by_HMA_winter) #check dataframe

mod5 = lm(formula= logit_poales ~ RAP_total_GF, data= graminoids_by_HMA_winter)
summary(mod5)

#### Biomass
cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$RAP_herb_biomass, method="spearman") #not significant p=0.1735 rho= -0.3714

mod6 = lm(formula= logit_poales ~ RAP_herb_biomass, data= graminoids_by_HMA_winter)
summary(mod6)

####Herd Buffered cover

View(RAP_and_poales_by_HMA_winter) #check dataframe

mod7 = lm(formula= logit_poales ~ mean_herb_cover, data= RAP_and_poales_by_HMA_winter)
summary(mod7)
#### Herd buffered biomass


cor.test(RAP_and_poales_by_HMA_winter$mean_herb_biomass, RAP_and_poales_by_HMA_winter$logit_poales, method="pearson")#not significant p=0.471 cor = -0.202

mod8 = lm(formula= logit_poales ~ mean_herb_biomass, data= RAP_and_poales_by_HMA_winter)
summary(mod8)

```


##Response transformed Graminoids, predictor cover transformed
```{r}

##Summer (no SB) 

#### Cover


View(graminoids_by_HMA_summer2) #check dataframe
mod9 = lm(formula= logit_poales ~ logit_cover, data= graminoids_by_HMA_summer2)
summary(mod9)

####Herd Buffered cover

View(RAP_and_poales_by_HMA_summer2) #check dataframe


mod10 = lm(formula= logit_poales ~ logit_cover, data= RAP_and_poales_by_HMA_summer2)
summary(mod10)

# Winter 

#### Cover

View(graminoids_by_HMA_winter) #check dataframe

mod11 = lm(formula= logit_poales ~ logit_cover, data= graminoids_by_HMA_winter)
summary(mod11)

####Herd Buffered cover
cor.test(RAP_and_poales_by_HMA_winter$logit_cover, RAP_and_poales_by_HMA_winter$logit_poales, method="spearman")#not significant p=0.3334 rho = -0.2679

View(RAP_and_poales_by_HMA_winter) #check dataframe

mod12 = lm(formula= logit_poales ~ logit_cover, data= RAP_and_poales_by_HMA_winter)
summary(mod12)
```


## Response BCS, predictor graminoids transformed
```{r}

## Summer

View(graminoids_by_HMA_summer2)
mod13 = lm(formula= mean_BCS ~ logit_poales, data= graminoids_by_HMA_summer2)
summary(mod13)

## Winter 

cor.test(graminoids_by_HMA_winter$logit_poales, graminoids_by_HMA_winter$mean_BCS, method="pearson") # not significant p=0.9322 cor = -0.02405

View(graminoids_by_HMA_winter)
mod14 = lm(formula= mean_BCS ~ logit_poales, data= graminoids_by_HMA_winter)
summary(mod14)


###### Summer graminoid and winter body condition #############

cor.test(graminoids_by_HMA_winter$summer_logit_poales, graminoids_by_HMA_winter$mean_BCS, method="spearman") # not significant p=0.8124 rho = -0.067857


View(graminoids_by_HMA_winter)
mod15 = lm(formula= mean_BCS ~ summer_logit_poales, data= graminoids_by_HMA_winter)
summary(mod15)

```

#Range model incl. cover and ecosystem
```{r}

#View (range_dif) #this is the values that calculate difference HMA average summer/winter

range_dif_df<- as.data.frame(range_dif) #make it a dataframe

View(range_dif_df) #check it

View(graminoids_by_HMA_winter) #view dataframe to pull values from 

range_dif_df$study.area<- graminoids_by_HMA_winter$study.area
range_dif_df$epa_eco<-as.factor(graminoids_by_HMA_winter$epa_eco)#add ecosystem as a factor
range_dif_df$RAP_total_GF<-graminoids_by_HMA_winter$RAP_total_GF #add the RAP herbaceous cover data
range_dif_df$herb_cover_group<- c("Low", "Medium", "Low", "Medium", "Low","Medium", "Medium", "High", "Medium", "Low", "Low", "High","High", "High","High")#create factor for cover grouping high, med,low
range_dif_df$herb_cover_group<-as.factor(range_dif_df$herb_cover_group) #make it a factor

View(range_dif_df) #check it again

#create some models

range_cover_lm<- lm(range_dif~ herb_cover_group, data=range_dif_df) #lm with cover factors
summary(range_cover_lm)
anova(range_cover_lm)

range_eco_lm<- lm(range_dif~ epa_eco, data=range_dif_df) #lm with ecosystem factor 
summary(range_eco_lm) 
anova(range_eco_lm)

range_cover_eco_lm<- lm(range_dif~ epa_eco + herb_cover_group, data=range_dif_df) #lm with cover and ecosystem factors 
summary(range_cover_eco_lm)
anova(range_cover_eco_lm)

range_cover_eco_int_lm<- lm(range_dif~ epa_eco + herb_cover_group + herb_cover_group:epa_eco, data=range_dif_df) #lm with cover and ecosystem factors with interaction of the two  
summary(range_cover_eco_int_lm)
anova(range_cover_eco_int_lm)





```

#Table of families by HMA
```{r}

plant_family_HMA_season<-plant_family_meta_wild #move to new dataframe

#make each family a numeric that has not already been made numberic
plant_family_HMA_season$Chenopodiaceae<-as.numeric(plant_family_HMA_season$Chenopodiaceae)
plant_family_HMA_season$Polygonaceae<-as.numeric(plant_family_HMA_season$Polygonaceae)
plant_family_HMA_season$Asteraceae<-as.numeric(plant_family_HMA_season$Asteraceae)
plant_family_HMA_season$Fabaceae<-as.numeric(plant_family_HMA_season$Fabaceae)
plant_family_HMA_season$Brassicaceae<-as.numeric(plant_family_HMA_season$Brassicaceae)
plant_family_HMA_season$Sarcobataceae<-as.numeric(plant_family_HMA_season$Sarcobataceae)
plant_family_HMA_season$Fagaceae<-as.numeric(plant_family_HMA_season$Fagaceae)
plant_family_HMA_season$Ephedraceae<-as.numeric(plant_family_HMA_season$Ephedraceae)
plant_family_HMA_season$Ranunculaceae<-as.numeric(plant_family_HMA_season$Ranunculaceae)
plant_family_HMA_season$Malvaceae<-as.numeric(plant_family_HMA_season$Malvaceae)
plant_family_HMA_season$Rosaceae<-as.numeric(plant_family_HMA_season$Rosaceae)
plant_family_HMA_season$Hydrophyllaceae<-as.numeric(plant_family_HMA_season$Hydrophyllaceae)
plant_family_HMA_season$Tamaricaceae<-as.numeric(plant_family_HMA_season$Tamaricaceae)
plant_family_HMA_season$Geraniaceae<-as.numeric(plant_family_HMA_season$Geraniaceae)
plant_family_HMA_season$Potamogetonaceae<-as.numeric(plant_family_HMA_season$Potamogetonaceae)
plant_family_HMA_season$Apiaceae<-as.numeric(plant_family_HMA_season$Apiaceae)
plant_family_HMA_season$Berberidaceae<-as.numeric(plant_family_HMA_season$Berberidaceae)
plant_family_HMA_season$Boraginaceae<-as.numeric(plant_family_HMA_season$Boraginaceae)
plant_family_HMA_season$Onagraceae<-as.numeric(plant_family_HMA_season$Onagraceae)
plant_family_HMA_season$Polemoniaceae<-as.numeric(plant_family_HMA_season$Polemoniaceae)
plant_family_HMA_season$Pinaceae<-as.numeric(plant_family_HMA_season$Pinaceae)
plant_family_HMA_season$Cupressaceae<-as.numeric(plant_family_HMA_season$Cupressaceae)
plant_family_HMA_season$Nyctaginaceae<-as.numeric(plant_family_HMA_season$Nyctaginaceae)
plant_family_HMA_season$Plantaginaceae<-as.numeric(plant_family_HMA_season$Plantaginaceae)
plant_family_HMA_season$Comandraceae<-as.numeric(plant_family_HMA_season$Comandraceae)
plant_family_HMA_season$Salicaceae<-as.numeric(plant_family_HMA_season$Salicaceae)
plant_family_HMA_season$Bignoniaceae<-as.numeric(plant_family_HMA_season$Bignoniaceae)
plant_family_HMA_season$Caryophyllaceae<-as.numeric(plant_family_HMA_season$Caryophyllaceae)
plant_family_HMA_season$Urticaceae<-as.numeric(plant_family_HMA_season$Urticaceae)
plant_family_HMA_season$Campanulaceae<-as.numeric(plant_family_HMA_season$Campanulaceae)
plant_family_HMA_season$Krameriaceae<-as.numeric(plant_family_HMA_season$Krameriaceae)
plant_family_HMA_season$Scrophulariaceae<-as.numeric(plant_family_HMA_season$Scrophulariaceae)
plant_family_HMA_season$Loasaceae<-as.numeric(plant_family_HMA_season$Loasaceae)
plant_family_HMA_season$Stachyuraceae<-as.numeric(plant_family_HMA_season$Stachyuraceae)
plant_family_HMA_season$Caprifoliaceae<-as.numeric(plant_family_HMA_season$Caprifoliaceae)
plant_family_HMA_season$Amblystegiaceae<-as.numeric(plant_family_HMA_season$Amblystegiaceae)
plant_family_HMA_season$Zygophyllaceae<-as.numeric(plant_family_HMA_season$Zygophyllaceae)
plant_family_HMA_season$Asparagaceae<-as.numeric(plant_family_HMA_season$Asparagaceae)
plant_family_HMA_season$Linaceae<-as.numeric(plant_family_HMA_season$Linaceae)
plant_family_HMA_season$Phrymaceae<-as.numeric(plant_family_HMA_season$Phrymaceae)
plant_family_HMA_season$Orobanchaceae<-as.numeric(plant_family_HMA_season$Orobanchaceae)
plant_family_HMA_season$Anacardiaceae<-as.numeric(plant_family_HMA_season$Anacardiaceae)
plant_family_HMA_season$Amaryllidaceae<-as.numeric(plant_family_HMA_season$Amaryllidaceae)
plant_family_HMA_season$Bryaceae<-as.numeric(plant_family_HMA_season$Bryaceae)
plant_family_HMA_season$Cucurbitaceae<-as.numeric(plant_family_HMA_season$Cucurbitaceae)
plant_family_HMA_season$Juncaginaceae<-as.numeric(plant_family_HMA_season$Juncaginaceae)
plant_family_HMA_season$Solanaceae<-as.numeric(plant_family_HMA_season$Solanaceae)

#export
write.csv(plant_family_HMA_season, "family_by_sample.csv")# write to csv dataframe for each sample

#get the HMA/season averages
plant_family_HMA_season_sd<-plant_family_HMA_season%>%   
  group_by(study.area, season)%>%
  summarise(Poaceae1=mean(Poaceae), sd_Poaceae=sd(Poaceae), Cyperaceae1=mean(Cyperaceae), sd_Cyperaceae=sd(Cyperaceae), Juncaceae1=mean(Juncaceae), sd_Juncaceae=sd(Juncaceae), Asteraceae1=mean(Asteraceae), sd_Asteraceae=sd(Asteraceae), Chenopodiaceae1=mean(Chenopodiaceae), sd_Chenopodiaceae=sd(Chenopodiaceae),Polygonaceae1=mean(Polygonaceae), sd_Polygonaceae=sd(Polygonaceae), Fabaceae1=mean(Fabaceae), sd_Fabaceae=sd(Fabaceae),Brassicaceae1=mean(Brassicaceae), sd_Brassicaceae=sd(Brassicaceae),Sarcobataceae1=mean(Sarcobataceae), sd_Sarcobataceae=sd(Sarcobataceae),Fagaceae1=mean(Fagaceae), sd_Fagaceae=sd(Fagaceae),Ephedraceae1=mean(Ephedraceae), sd_Ephedraceae=sd(Ephedraceae),Ranunculaceae1=mean(Ranunculaceae), sd_Ranunculaceae=sd(Ranunculaceae),Malvaceae1=mean(Malvaceae), sd_Malvaceae=sd(Malvaceae),Rosaceae1=mean(Rosaceae), sd_Rosaceae=sd(Rosaceae),Hydrophyllaceae1=mean(Hydrophyllaceae), sd_Hydrophyllaceae=sd(Hydrophyllaceae), Tamaricaceae1=mean(Tamaricaceae), sd_Tamaricaceae=sd(Tamaricaceae), Geraniaceae1=mean(Geraniaceae), sd_Geraniaceae=sd(Geraniaceae),Potamogetonaceae1=mean(Potamogetonaceae), sd_Potamogetonaceae=sd(Potamogetonaceae), Apiaceae1=mean(Apiaceae), sd_Apiaceae=sd(Apiaceae), Berberidaceae1=mean(Berberidaceae), sd_Berberidaceae=sd(Berberidaceae), Boraginaceae1=mean(Boraginaceae), sd_Boraginaceae=sd(Boraginaceae), Onagraceae1=mean(Onagraceae), sd_Onagraceae=sd(Onagraceae), Polemoniaceae1=mean(Polemoniaceae), sd_Polemoniaceae=sd(Polemoniaceae), Pinaceae1=mean(Pinaceae), sd_Pinaceae=sd(Pinaceae), Cupressaceae1=mean(Cupressaceae), sd_Cupressaceae=sd(Cupressaceae), Nyctaginaceae1=mean(Nyctaginaceae), sd_Nyctaginaceae=sd(Nyctaginaceae), Plantaginaceae1=mean(Plantaginaceae), sd_Plantaginaceae=sd(Plantaginaceae), Comandraceae1=mean(Comandraceae), sd_Comandraceae=sd(Comandraceae),  Salicaceae1=mean(Salicaceae), sd_Salicaceae=sd(Salicaceae), Bignoniaceae1=mean(Bignoniaceae), sd_Bignoniaceae=sd(Bignoniaceae), Caryophyllaceae1=mean(Caryophyllaceae), sd_Caryophyllaceae=sd(Caryophyllaceae), Urticaceae1=mean(Urticaceae), sd_Urticaceae=sd(Urticaceae), Campanulaceae1=mean(Campanulaceae), sd_Campanulaceae=sd(Campanulaceae),  Krameriaceae1=mean(Krameriaceae), sd_Krameriaceae=sd(Krameriaceae), Scrophulariaceae1=mean(Scrophulariaceae), sd_Scrophulariaceae=sd(Scrophulariaceae), Loasaceae1=mean(Loasaceae), sd_Loasaceae=sd(Loasaceae), Stachyuraceae1=mean(Stachyuraceae), sd_Stachyuraceae=sd(Stachyuraceae), Caprifoliaceae1=mean(Caprifoliaceae), sd_Caprifoliaceae=sd(Caprifoliaceae), Amblystegiaceae1=mean(Amblystegiaceae), sd_Amblystegiaceae=sd(Amblystegiaceae), Zygophyllaceae1=mean(Zygophyllaceae), sd_Zygophyllaceae=sd(Zygophyllaceae), Asparagaceae1=mean(Asparagaceae), sd_Asparagaceae=sd(Asparagaceae), Linaceae1=mean(Linaceae), sd_Linaceae=sd(Linaceae), Phrymaceae1=mean(Phrymaceae), sd_Phrymaceae=sd(Phrymaceae), Orobanchaceae1=mean(Orobanchaceae), sd_Orobanchaceae=sd(Orobanchaceae), Anacardiaceae1=mean(Anacardiaceae), sd_Anacardiaceae=sd(Anacardiaceae), Amaryllidaceae1=mean(Amaryllidaceae), sd_Amaryllidaceae=sd(Amaryllidaceae), Bryaceae1=mean(Bryaceae), sd_Bryaceae=sd(Bryaceae), Cucurbitaceae1=mean(Cucurbitaceae), sd_Cucurbitaceae=sd(Cucurbitaceae), Juncaginaceae1=mean(Juncaginaceae), sd_Juncaginaceae=sd(Juncaginaceae), Solanaceae1=mean(Solanaceae), sd_Solanaceae=sd(Solanaceae))

View(plant_family_HMA_season_sd)

#export
write.csv(plant_family_HMA_season_sd, "family_by_HMA.csv")# write to csv dataframe for each HMA/season

#HMA averages without the SD
plant_family_HMA_season_no_sd<-plant_family_HMA_season%>%   
  group_by(study.area, season)%>%
  summarise(Poaceae1=mean(Poaceae), Cyperaceae1=mean(Cyperaceae),  Juncaceae1=mean(Juncaceae),  Asteraceae1=mean(Asteraceae),  Chenopodiaceae1=mean(Chenopodiaceae), Polygonaceae1=mean(Polygonaceae),  Fabaceae1=mean(Fabaceae), Brassicaceae1=mean(Brassicaceae), Sarcobataceae1=mean(Sarcobataceae), Fagaceae1=mean(Fagaceae), Ephedraceae1=mean(Ephedraceae), Ranunculaceae1=mean(Ranunculaceae), Malvaceae1=mean(Malvaceae), Rosaceae1=mean(Rosaceae), Hydrophyllaceae1=mean(Hydrophyllaceae), Tamaricaceae1=mean(Tamaricaceae),  Geraniaceae1=mean(Geraniaceae), Potamogetonaceae1=mean(Potamogetonaceae), Apiaceae1=mean(Apiaceae),  Berberidaceae1=mean(Berberidaceae),  Boraginaceae1=mean(Boraginaceae),  Onagraceae1=mean(Onagraceae),  Polemoniaceae1=mean(Polemoniaceae),  Pinaceae1=mean(Pinaceae),  Cupressaceae1=mean(Cupressaceae), Nyctaginaceae1=mean(Nyctaginaceae),  Plantaginaceae1=mean(Plantaginaceae), Comandraceae1=mean(Comandraceae),  Salicaceae1=mean(Salicaceae),  Bignoniaceae1=mean(Bignoniaceae),  Caryophyllaceae1=mean(Caryophyllaceae), Urticaceae1=mean(Urticaceae),  Campanulaceae1=mean(Campanulaceae),   Krameriaceae1=mean(Krameriaceae),  Scrophulariaceae1=mean(Scrophulariaceae),  Loasaceae1=mean(Loasaceae), Stachyuraceae1=mean(Stachyuraceae),  Caprifoliaceae1=mean(Caprifoliaceae),  Amblystegiaceae1=mean(Amblystegiaceae), Zygophyllaceae1=mean(Zygophyllaceae),  Asparagaceae1=mean(Asparagaceae), Linaceae1=mean(Linaceae), Phrymaceae1=mean(Phrymaceae), Orobanchaceae1=mean(Orobanchaceae),  Anacardiaceae1=mean(Anacardiaceae),  Amaryllidaceae1=mean(Amaryllidaceae),  Bryaceae1=mean(Bryaceae),  Cucurbitaceae1=mean(Cucurbitaceae),  Juncaginaceae1=mean(Juncaginaceae),  Solanaceae1=mean(Solanaceae))

View(plant_family_HMA_season_no_sd)
#export
write.csv(plant_family_HMA_season_no_sd, "family_by_HMA_noSD.csv")# write to csv dataframe for each HMA/season
```
#Asteraceae/Chenopodiaceae comparisons
```{r}
#calculate range-wide SD/ mean (including SB in summer)

mean(subset(plant_family_HMA_season_no_sd, season == "winter")$Asteraceae1) # 15.05261
sd(subset(plant_family_HMA_season_no_sd, season == "winter")$Asteraceae1) # 15.7874

mean(subset(plant_family_HMA_season_no_sd, season == "summer")$Asteraceae1) # 6.03408
sd(subset(plant_family_HMA_season_no_sd, season == "summer")$Asteraceae1) # 5.794897

mean(subset(plant_family_HMA_season_no_sd, season == "winter")$Chenopodiaceae1)# 14.53212
sd(subset(plant_family_HMA_season_no_sd, season == "winter")$Chenopodiaceae1) # 15.22876

mean(subset(plant_family_HMA_season_no_sd, season == "summer")$Chenopodiaceae1) # 3.9063
sd(subset(plant_family_HMA_season_no_sd, season == "summer")$Chenopodiaceae1) # 8.1305


#t-tests for Chenopodiaceae and Asteraceae
range_cheno_summer<-(subset(plant_family_HMA_season_no_sd, study.area != "SB"))
range_cheno_summer<-(subset(range_cheno_summer, season == "summer")$Chenopodiaceae1) 
range_cheno_winter<-(subset(plant_family_HMA_season_no_sd, season == "winter")$Chenopodiaceae1) 

set.seed(223)
boot.t.test(x=range_cheno_summer, y=range_cheno_winter, paired=TRUE) #paired bootstrap 0.0144, paired traditional p=0.037

range_aster_summer<-(subset(plant_family_HMA_season_no_sd, study.area != "SB"))
range_aster_summer<-(subset(range_aster_summer, season == "summer")$Asteraceae1) 
range_aster_winter<-(subset(plant_family_HMA_season_no_sd, season == "winter")$Asteraceae1) 

set.seed(223)
boot.t.test(x=range_aster_summer, y=range_aster_winter, paired=TRUE) #paired bootstrap 0.0096, paired traditional p=0.028


#calculate range-wide SD/ mean (excluding SB in summer)
mean(range_cheno_summer) # 4.06774
sd(range_cheno_summer)# 8.38929

mean(range_aster_summer)# 5.69809
sd(range_aster_summer) # 5.834804

```
mean(range_aster_summer)# 5.69809
sd(range_aster_summer) # 5.834804

```
